---
title: "pH Calculation Comparison AE17 Experiment"
author: "Alan Downey-Wall"
date: "7/29/2020"
output: github_document
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/home/downeyam/Github/AE17_Cvirginica_MolecularResponse')
```

**Overview** : The aim of this document is to compare the effect of different pH scales and evaluate the impact of pH scale on the results and interpretation of the AE17 adult exposure experiment.

**Reason** : The reason for the this exploration was due to the fact that the EPF and seawater pH probes were measured with different instruments, liquid junction and solid state, respectively. The offsets among measuring instruments was not properly corrected, so below we examine the impact of including this correction, specifically examining the corrected SW and EPF pHs on the `Total`, `Seawater`, and `NBS` scales.

```{r include=FALSE}
##Libraries and Misc. Scripts
library(mgcv)
library(dplyr)
library(car)
library(lme4)
library(lmerTest)
library(factoextra)
library(kableExtra)
library(multcomp)
library(multcompView)
library(ggplot2)
library(cowplot)
library("RColorBrewer")
pal <- brewer.pal(n = 12, name = 'Paired')
col_perm <- c(pal[1:2],pal[5:6],pal[12])
# Located in src Analysis/Phenotype folder, will need to set full working directory or setwd()
#setwd("/home/downeyam/Github/AE17_Cvirginica_MolecularResponse")
source("src/Accessory/basicR_functions.R")
```

```{r echo=FALSE}
# Sample Data
epf_exp <- read.csv("data/Phenotype/CompletePhenotype_final2020.csv",stringsAsFactors = FALSE)
epf_exp$pCO2_fac <- as.factor(epf_exp$pCO2)
epf_exp$Timepoint_fac <- as.factor(epf_exp$Timepoint)
# Weekly Water Chem
wc <- read.delim("data/water_chem/AE17_weeklyExposure_final2020.csv",sep=",",
                 stringsAsFactors = FALSE)
```

## Data 
```{r echo=FALSE}
col <- c("ID","Date","Timepoint","TankID",
         "EPF_pH_raw","EPF_NBS_pH","EPF_Total_pH",
         "pH_NBS_2W","pH_scaleFree_2W","pH_Total_2W")
kab_table <-subset(epf_exp,select=col)

kable(kab_table,align="c" ) %>%
  kable_styling() %>%
  scroll_box(width = "110%", height = "400px")
```

```{r echo=FALSE}
# Stats summary function
# Data should include
#   Col 1 : EPF pH measure (pick a scale)
#   Col 2 : SW (2W) pH (pick a scale)
#   Col 3 : pCO2 (factor)
#   Col 4 : Time point (factor)
#   Col 5 : TankID (factor)
statSummary <- function(x=NULL){
  col <- c("epH","spH","t","tp","tank")
  colnames(x) <- col
  # Calculate delta epf
  x$epfd <- x$epH - x$spH
  
  # Linear mixed model
  out_epf <- lmer(epH~t*tp + (1|tank),data=x)
  anova_epf <- anova(out_epf)
  out_epfd <- lmer(epfd~t*tp + (1|tank),data=x)
  anova_epfd <- anova(out_epfd)
  
  ## Planned Comparisons (control vs. oa treatments, both EPF pH and delta pH)
  group <- paste0(x$t,":",x$tp)
  mod <- model.matrix(out_epf)
  agg<- aggregate(mod~group,FUN=mean)
  rownames(agg) <- agg$group
  agg <- agg[,-1]
  lc2 <- as.matrix(agg)
  k <- rbind("400v900_1"=lc2["900:1",]    -lc2["400:1",],
           "400v2800_1"=lc2["2800:1",]  -lc2["400:1",],
           "400v900_2"=lc2["900:2",]    -lc2["400:2",],
           "400v2800_2"=lc2["2800:2",]  -lc2["400:2",],
           "400v900_9"=lc2["900:9",]    -lc2["400:9",],
           "400v2800_9"=lc2["2800:9",]  -lc2["400:9",],
           "400v900_22"=lc2["900:22",]  -lc2["400:22",],
           "400v2800_22"=lc2["2800:22",]-lc2["400:22",],
           "400v900_50"=lc2["900:50",]  -lc2["400:50",],
           "400v2800_50"=lc2["2800:50",]-lc2["400:50",],
           "400v900_80"=lc2["900:80",]  -lc2["400:80",],
           "400v2800_80"=lc2["2800:80",]-lc2["400:80",])
  out_epf_ph <- summary(glht(out_epf,linfct=k),adjusted(type = "fdr"))
  out_epfd_ph <- summary(glht(out_epfd,linfct=k),adjusted(type = "fdr"))
  
  # T.test to look at delta pH relative to environment
  t_epfd<-NULL
  for(i in unique(x$tp)){
    for(j in unique(x$t)){
      t_epfd<-c(t_epfd,t.test(x$epfd[x$tp == i & x$t == j],mu = 0)$p.value)
    }
  }
  temp <- expand.grid(unique(x$t),unique(x$tp))
  t_df <- data.frame(Timepoint=temp[,2],pCO2=temp[,1],pValue=t_epfd)
  t_df$pValue_corr <- p.adjust(t_df$pValue,method = "BH")
  t_df$Sig <- ifelse(t_df$pValue_corr < 0.05,"TRUE","FALSE")
  
  # Returns final analyses as a list
  final <- list(anova.epfpH=anova_epf,
                anova.deltapH=anova_epfd,
                pc.epfpH=out_epf_ph,
                pc.epfdpH=out_epfd_ph,
                T.values=t_df)
  return <- final
}

```

```{r echo=FALSE}
## Analysis

### Rerunning linear models with differen pH scales  

##Total pH scale
tot <- data.frame(epf_exp$EPF_Total_pH,epf_exp$pH_Total_2W,
                  epf_exp$pCO2_fac,epf_exp$Timepoint_fac,
                  epf_exp$TankID)
tot_stats <- statSummary(tot)
##NBS pH scale
nbs <- data.frame(epf_exp$EPF_NBS_pH,epf_exp$pH_NBS_2W,
                  epf_exp$pCO2_fac,epf_exp$Timepoint_fac,
                  epf_exp$TankID)
nbs_stats <- statSummary(nbs)
##Original (uncorrected) pH scale
or <- data.frame(epf_exp$EPF_pH_raw,epf_exp$pH_scaleFree_2W,
                  epf_exp$pCO2_fac,epf_exp$Timepoint_fac,
                  epf_exp$TankID)
or_stats <- statSummary(or)
```

# Results

## Plot comparing impact of adjustment on EPF and SW
```{r echo=FALSE}
# EPF
EPFpH <- c(epf_exp$EPF_NBS_pH-epf_exp$EPF_pH_raw,
        epf_exp$EPF_Total_pH-epf_exp$EPF_pH_raw)
# Seawater
SWpH <- c(epf_exp$pH_NBS_2W-epf_exp$pH_scaleFree_2W,
        epf_exp$pH_Total_2W-epf_exp$pH_scaleFree_2W)
Scale <- sort(rep(c("Total","NBS"),times=nrow(epf_exp)))
epf_long <- data.frame(type="EPF",Scale,pH_diff=EPFpH)
sw_long <- data.frame(type="SW",Scale,pH_diff=SWpH)
pH_long <- rbind(epf_long,sw_long)

p1 <- ggplot(pH_long,aes(x=Scale,y=pH_diff,colour=type)) + 
  geom_hline(yintercept=0) +
  geom_boxplot() +
  theme_cowplot() +
  labs(y="pH change from scale-free (Current)")
p1
```
Y-axis measured as the difference in pH from the corrected pH values from the original pH.

## Statistics Tables

### EPF pH

```{r echo=FALSE}
min_col <- rep(c("F","Pvalue"),times=3)
lmer_stats <- cbind(or_stats$anova.epfpH$`F value`,
or_stats$anova.epfpH$`Pr(>F)`,
nbs_stats$anova.epfpH$`F value`,
nbs_stats$anova.epfpH$`Pr(>F)`,
tot_stats$anova.epfpH$`F value`,
tot_stats$anova.epfpH$`Pr(>F)`)
lmer_stats <- data.frame(Effect=c("Treatment","Timepoint","Interaction"),lmer_stats)
colnames(lmer_stats) <- c("",min_col)

kable(lmer_stats,digits = 3, caption="EPF pH - Overall Linear Mixed Effects model") %>%
  kable_styling(full_width = TRUE) %>%
  add_header_above(c(" ", "Original" = 2, "NBS" = 2,"Total"=2))
```
  
```{r echo=FALSE}
pc_stats <- cbind(or_stats$pc.epfpH$test$pvalues,
nbs_stats$pc.epfpH$test$pvalues,
tot_stats$pc.epfpH$test$pvalues)
pc_stats <- data.frame(pc_stats)
colnames(pc_stats) <- c("Original P","NBS P","Total P")

kable(pc_stats,digits = 3, caption="EPF pH - Planned comparisons") %>%
  kable_styling()
```

### DELTA pH

```{r echo=FALSE}
min_col <- rep(c("F","Pvalue"),times=3)
lmer_stats <- cbind(or_stats$anova.epfpH$`F value`,
or_stats$anova.deltapH$`Pr(>F)`,
nbs_stats$anova.deltapH$`F value`,
nbs_stats$anova.deltapH$`Pr(>F)`,
tot_stats$anova.deltapH$`F value`,
tot_stats$anova.deltapH$`Pr(>F)`)
lmer_stats <- data.frame(Effect=c("Treatment","Timepoint","Interaction"),lmer_stats)
colnames(lmer_stats) <- c("",min_col)

kable(lmer_stats,digits = 3, caption="Delta EPF pH - Overall Linear Mixed Effects model") %>%
  kable_styling(full_width = TRUE) %>%
  add_header_above(c(" ", "Original" = 2, "NBS" = 2,"Total"=2))
```

```{r echo=FALSE}
pc_stats <- cbind(or_stats$pc.epfdpH$test$pvalues,
nbs_stats$pc.epfdpH$test$pvalues,
tot_stats$pc.epfdpH$test$pvalues)
pc_stats <- data.frame(pc_stats)
colnames(pc_stats) <- c("Original P","NBS P","Total P")

kable(pc_stats,digits = 3, caption="Delta EPF pH - Planned comparisons")%>%
  kable_styling()
```

```{r echo=FALSE}
t_stats <- cbind(or_stats$T.values[,c(4,5)],
nbs_stats$T.values[,c(4,5)],
tot_stats$T.values[,c(4,5)])
t_stats <- data.frame(x=or_stats$T.values[,c(1,2)],t_stats)
min_col <- c("Timepoint","pCO2","Pvalue1","Sig1",
             "Pvalue2","Sig2",
             "Pvalue3","Sig3")
colnames(t_stats) <- min_col
t_stats <- arrange(t_stats,pCO2,Timepoint)
t_stats %>%
  mutate(
    Sig1 = cell_spec(Sig1,background =ifelse(Sig1==TRUE, "lightgreen", "white")),
    Sig2 = cell_spec(Sig2,background =ifelse(Sig2==TRUE, "lightgreen", "white")),
    Sig3 = cell_spec(Sig3,background =ifelse(Sig3==TRUE, "lightgreen", "white"))
  )%>%
kable(digits = 3,align="c", caption="Delta EPF pH - T tests ",
      escape = F, booktabs = T, linesep = "") %>%
  kable_styling()%>%
  add_header_above(c(" "=2, "Original" = 2, "NBS" = 2,"Total"=2))
```

## Figure

```{r echo=FALSE}
# Function for creating standard plot used in manuscript
plotStandard <- function(x=NULL,c_mean=NULL,oa_900_mean=NULL,oa_2800_mean=NULL){
  col <- c("epH","spH","t","tp","tank")
  colnames(x) <- col
  # Calculate delta epf
  x$epfd <- x$epH - x$spH

  # Code for setting the plotting space for the two panels (plus middle section for     shared legend)
  m <- matrix(c(1,2,3),nrow = 3,ncol = 1,byrow = TRUE)
  layout(mat = m,heights = c(0.5,0.05,0.5),widths = c(.1))
  par(mar = c(5,5,3,2))
  
  #### Panel A - measured EPF pH ####
  ### Summarize measured EPF pH for plotting
  ## Take the means for each timepoint and treatment level
  treatSeq_means <- aggregate(epH~tp+t,x,FUN=mean)
  treatSeq_means$tp[treatSeq_means$tp == 2] <- 4
  treat_means_ctrl <- treatSeq_means[treatSeq_means$t == "400",]
  treat_means_oa_900 <- treatSeq_means[treatSeq_means$t == "900",]
  treat_means_oa_2800 <- treatSeq_means[treatSeq_means$t == "2800",]
  # Stagger the timepoints for 900 and 2800 treatments so they can be seen better on plot
  treat_means_oa_900$timepoint <- treat_means_oa_900$tp + 1
  treat_means_oa_2800$timepoint <- treat_means_oa_2800$tp + 2
  # Manually alter timepoint 2 placement along x-axis for visual clarity
  treat_means_oa_900$tp <- treat_means_oa_900$tp + 1
  treat_means_oa_2800$tp <- treat_means_oa_2800$tp +2
  ## Take the standard error (SE) for each timepoint and treatment level
  treatSeq_SE <- aggregate(epH~tp+t,x,FUN=se)
  treat_SE_ctrl <- treatSeq_SE[treatSeq_SE$t == "400",]$epH
  treat_SE_oa_900 <- treatSeq_SE[treatSeq_SE$t == "900",]$epH
  treat_SE_oa_2800 <- treatSeq_SE[treatSeq_SE$t == "2800",]$epH
  
  # Base plot with 400
  bp <- plot(treat_means_ctrl$epH~treat_means_ctrl$tp,
             ylab=expression(pH[EPF]~(NBS)),xlab="Time (Days)",
             col=col_perm[2],pch=16,cex=2,cex.axis = 1.5,cex.lab=1.5,
             ylim = c(7.0,8.25),xlim=c(0,82),bty="n")
  
  # Mean environment lines underneath other lines    
  abline(h=c_mean,col=col_perm[2],lty=2) # control
  abline(h=oa_900_mean,col=col_perm[5],lty=2) # 900
  abline(h=oa_2800_mean,col=col_perm[4],lty=2) # 2800
  
  # 400 EPF lines
  lines(treat_means_ctrl$epH~treat_means_ctrl$tp,
        col="lightblue4")
  arrows(x0 = treat_means_ctrl$tp,
         x1 = treat_means_ctrl$tp,
         y0 = c(treat_means_ctrl$epH - treat_SE_ctrl),
         y1 = treat_means_ctrl$epH + treat_SE_ctrl,
         angle = 90, len = 0.05,
         code = 3, xpd = NA, lwd = 2,
         col=col_perm[2])
  # 900 EPF lines
  points(treat_means_oa_900$epH~treat_means_oa_900$tp,
         col=col_perm[5],pch=15,cex=2)
  lines(treat_means_oa_900$epH~treat_means_oa_900$tp,
        col=col_perm[5])
  arrows(x0 = treat_means_oa_900$tp,
         x1 = treat_means_oa_900$tp,
         y0 = c(treat_means_oa_900$epH - treat_SE_ctrl),
         y1 = treat_means_oa_900$epH + treat_SE_ctrl,
         angle = 90, len = 0.05,
         code = 3, xpd = NA, lwd = 2,
         col=col_perm[5])
  # 2800 EPF lines 
  points(treat_means_oa_2800$epH~treat_means_oa_2800$tp,
         col=col_perm[4],pch=17,cex=2)
  lines(treat_means_oa_2800$epH~treat_means_oa_2800$tp,
        col=col_perm[4])
  arrows(x0 = treat_means_oa_2800$tp,
         x1 = treat_means_oa_2800$tp,
         y0 = c(treat_means_oa_2800$epH - treat_SE_oa_2800),
         y1 = treat_means_oa_2800$epH + treat_SE_oa_2800,
         angle = 90, len = 0.05, code = 3, xpd = NA, lwd = 2,
         col=col_perm[4])
  
  #### Legend #####
  par(mar = c(0,0,0,0))
  plot(1, type = "n", axes=FALSE, xlab="", ylab="")
  legend("bottom",horiz = TRUE,
         legend = c("Control","Moderate OA","High OA"),
         col = c(col_perm[2],col_perm[5],col_perm[4]),
         pch=c(16,15,17),
         cex = 1.5,
         lwd = 2,
         bty = "n")
  
  #### Panel B - diff EPF #### 
  par(mar = c(5,5,3,2))
  ## Take the means for each timepoint and treatment level
  treatSeq_means <- aggregate(epfd~tp+t,x,FUN=mean)
  treatSeq_means$tp[treatSeq_means$tp == 2] <- 4 
  treat_means_ctrl <- treatSeq_means[treatSeq_means$t == "400",]
  treat_means_oa_900 <- treatSeq_means[treatSeq_means$t == "900",]
  treat_means_oa_2800 <- treatSeq_means[treatSeq_means$t == "2800",]
  # Stagger the timepoints for 900 and 2800 treatments so they can be seen better on plot
  # Manually alter timepoint 2 placement along x-axis for visual clarity
  treat_means_oa_900$tp <- treat_means_oa_900$tp+ 1
  treat_means_oa_2800$tp <- treat_means_oa_2800$tp+ 2
  ## Take the standard error (se) for each timepoint and treatment level
  treatSeq_SE <- aggregate(epfd~tp+t,x,FUN=se)
  treat_SE_ctrl <- treatSeq_SE[treatSeq_SE$t == "400",]$epfd
  treat_SE_oa_900 <- treatSeq_SE[treatSeq_SE$t == "900",]$epfd
  treat_SE_oa_2800 <- treatSeq_SE[treatSeq_SE$t == "2800",]$epfd
  
  # Base plot with 400
  bp <- plot(treat_means_ctrl$epfd~treat_means_ctrl$tp,
             ylab=expression(paste(Delta," pH (NBS)")),
             xlab="Time (Days)",pch=16,
             col=col_perm[2],cex=2,cex.axis = 1.5,cex.lab=1.5,
             ylim = c(-1.2,1.1),xlim=c(0,82),bty="n")
  # Mean environment lines underneath other lines    
  abline(h=0,col="black",lty=2) # control
  # 400 EPF lines
  lines(treat_means_ctrl$epfd~treat_means_ctrl$tp,
        col=col_perm[2])
  arrows(x0 = treat_means_ctrl$tp,
         x1 = treat_means_ctrl$tp,
         y0 = c(treat_means_ctrl$epfd - treat_SE_ctrl),
         y1 = treat_means_ctrl$epfd + treat_SE_ctrl,
         angle = 90, len = 0.05,
         code = 3, xpd = NA, lwd = 2,
         col=col_perm[2])
  # 900 EPF lines
  points(treat_means_oa_900$epfd~treat_means_oa_900$tp,
         col=col_perm[5],pch=15,cex=2)
  lines(treat_means_oa_900$epfd~treat_means_oa_900$tp,
        col=col_perm[5])
  arrows(x0 = treat_means_oa_900$tp,
         x1 = treat_means_oa_900$tp,
         y0 = c(treat_means_oa_900$epfd - treat_SE_oa_900),
         y1 = treat_means_oa_900$epfd + treat_SE_oa_900,
         angle = 90, len = 0.05,
         code = 3, xpd = NA, lwd = 2,
         col=col_perm[5])
  # 2800 EPF lines 
  points(treat_means_oa_2800$epfd~treat_means_oa_2800$tp,
         col=col_perm[4],pch=17,cex=2)
  lines(treat_means_oa_2800$epfd~treat_means_oa_2800$tp,
        col=col_perm[4])
  arrows(x0 = treat_means_oa_2800$tp,
         x1 = treat_means_oa_2800$tp,
         y0 = c(treat_means_oa_2800$epfd - treat_SE_oa_2800),
         y1 = treat_means_oa_2800$epfd + treat_SE_oa_2800,
         angle = 90, len = 0.05, code = 3, xpd = NA, lwd = 2,
         col=col_perm[4])
}
```
  
### Original pH scale 
```{r fig.width=10,fig.height=8}
or <- data.frame(epf_exp$EPF_pH_raw,epf_exp$pH_scaleFree_2W,
                  epf_exp$pCO2,epf_exp$Timepoint,
                  epf_exp$TankID)
#Original
c_mean_O <- mean(wc$pH_scaleFree[wc$PCO2 == 550])
oa_900_mean_O <- mean(wc$pH_scaleFree[wc$PCO2 == 1000])
oa_2800_mean_O <- mean(wc$pH_scaleFree[wc$PCO2 == 2800])

plotStandard(or,c_mean_O,oa_900_mean_O,oa_2800_mean_O)
```

### NBS pH scale 
```{r fig.width=10,fig.height=8}
nbs <- data.frame(epf_exp$EPF_NBS_pH,epf_exp$pH_NBS_2W,
                  epf_exp$pCO2,epf_exp$Timepoint,
                  epf_exp$TankID)
#NBS
c_mean_N <- mean(wc$pH_NBS[wc$PCO2 == 550])
oa_900_mean_N <- mean(wc$pH_NBS[wc$PCO2 == 1000])
oa_2800_mean_N <- mean(wc$pH_NBS[wc$PCO2 == 2800])

plotStandard(nbs,c_mean_N,oa_900_mean_N,oa_2800_mean_N)
```

### Total pH scale 
```{r fig.width=10,fig.height=8}
tot <- data.frame(epf_exp$EPF_Total_pH,epf_exp$pH_Total_2W,
                  epf_exp$pCO2,epf_exp$Timepoint,
                  epf_exp$TankID)
#Total
c_mean_T <- mean(wc$pH_Total[wc$PCO2 == 550])
oa_900_mean_T <- mean(wc$pH_Total[wc$PCO2 == 1000])
oa_2800_mean_T <- mean(wc$pH_Total[wc$PCO2 == 2800])

plotStandard(tot,c_mean_T,oa_900_mean_T,oa_2800_mean_T)
```

## Conclusions  
  
**General Conclusion**
The correction appears to really only impact our delta EPF pH results, specifically when we are looking at the EPF pH compared to seawater. Both corrections significantly increased the seawater pH relative to the EPF pH. Overall, this led to EPF pH (regardless of treatment) being lower than the pH of seawater (within the respective treatment).  

**Difference between corrections**
The difference between `Total` and `NBS` scales is subtle, but the `NBS` led to a greater overall upward shift in pH, but this increase seems to be greatest in seawater pH, while Total seems to be an intermediate between NBS and the original calculation. Both corrections produce similar results, although we do not see a significant increase in EPF pH relative to seawater at day 9 when we use the `NBS` correction but we do see a significant increase with the `Total` correction. Given that we specifically mention this results as possible evidence of short term moderation of EPF pH, this does seem significant.

**The net negative trend in delta EPF pH**
The major difference that results from this correction compared to the literature is that we see a net negative delta pH. I think going back to some of the original Crenshaw work, perhaps this makes sense if the EPF continues to be a dumping ground for respiration / metabolism, but i am open to suggestions. I am also activity reviewing this literature again.
