---
title: "02 Salmon Pipeline"
author: "adowneywall"
date: "6/11/2019"
output: 
  html_document:
    keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DESeq2)
library(ashr)
library(vegan)
library(sf)
library(adegenet)
library(ggplot2)
```

## **Script Description**  

**Brief Overview** : This script performs a series of basic multivariate approaches to explore the whole genome expression profiles of the 24 oyster RNAseq samples. In partiular this includes:
* A permanova (implemented in the package vegan, called adonis)  
* An RDA (implemented in vegan)
* A DAPC with a focus on treatment(implemented in adegenet)

## **Data**  

Steps:  
* Read in dataframe with metadata for each samples including, treatment, time, population, lane of sequencing, and variable that contains each unique combination level between treatment and time.  
* Read in gene (or transcript) abundance matrix generated from the '01_salmon_countMatrix.Rmd' script.  
* Remove uninformative genes from count matrix.  

```{r}
## Meta Data for the Model
model<-read.csv("/home/downeyam/Github/2017OAExp_Oysters/input_files/RNA/metadata_cvirginica_rna_meta.txt", header=TRUE)
model$Treatment <- as.factor(model$treatment)
model$Time <- as.character(model$timepoint)
model$Time[model$Time == "3"] <- "09"
model$Time[model$Time == "6"] <- "80"
model$Time <- as.factor(model$Time)
model$Pop <- as.factor(model$population)
model$Lane <- as.factor(model$lane)
model$SFV <-  interaction(model$Time,model$Treatment) # Creates single factor variable for combination of time and treatment

gc <- readRDS("/home/downeyam/Github/2017OAExp_Oysters/input_files/RNA/salmon_pipeline/run20190610/geneMatrixCounts_default.Rdata")
#gc <- gc[-c(1:205),] #removing some no LOC genes (most trna)
```

## PERMANOVA (implements using adonis from vegan package)

* Testing for statistic differential between treatment, time, and treatment:time  

```{r}
(out <- adonis(t(gc)~Treatment*Time+Pop+Lane,data=model,permutations = 5000))
```

## Plotting data with RDA  

### Plot in multivariate space with RDA (treatment and time)
```{r}
prin_comp<-rda(t(gc), scale=FALSE)
sum_pri <- summary(prin_comp)
pca_scores<-scores(prin_comp)

pca <- prcomp(t(gc))
eigs <- pca$sdev^2

head(sum_pri$species)
color_comb <- c("deepskyblue2","blue4","firebrick1","darkred") # colors for population 
model$colors <- "" 
model$colors[model$SFV == unique(model$SFV)[1]] <-  color_comb[2]
model$colors[model$SFV == unique(model$SFV)[2]] <-  color_comb[1]
model$colors[model$SFV == unique(model$SFV)[3]] <-  color_comb[4]
model$colors[model$SFV == unique(model$SFV)[4]] <-  color_comb[3]

ordiplot(prin_comp,type="n",
         xlab=paste0("PC1 ",round(eigs[1] / sum(eigs)*100,1),"% Variance Explained)"),
         ylab=paste0("PC2 ",round(eigs[2] / sum(eigs)*100,1),"% Variance Explained)"))
orditorp(prin_comp,display="sites",labels = FALSE,col=model$colors,cex = 2,pch = 16,)
#ordiellipse(prin_comp,model$SFV,conf=0.90,col = color_comb,lwd = 3)
ordispider(prin_comp,model$SFV,col = color_comb,lwd=2.5)
legend(x=100,y=150,legend = c("Day09_Control","Day80_Control","Day09_OA","Day80_OA"),pch = 16,col=color_comb,xpd = .25)
text(x = -538 ,y = -300, paste0("Adonis P_Treatment = ",out$aov.tab$`Pr(>F)`[1],"*"))
text(x = -550 ,y = -360, paste0("Adonis P_Duration = ",out$aov.tab$`Pr(>F)`[2],"*"))
text(x = -540 ,y = -420, paste0("Adonis P_Interaction = ",out$aov.tab$`Pr(>F)`[5]))
```

### Plot colored by population  
```{r}
ordiplot(prin_comp,type="n",
         xlab=paste0("PC1 ",round(eigs[1] / sum(eigs)*100,1),"% Variance Explained)"),
         ylab=paste0("PC2 ",round(eigs[2] / sum(eigs)*100,1),"% Variance Explained)"))
orditorp(prin_comp,display="sites",labels = FALSE,col=model$Pop,cex = 2,pch = 16)
legend(x=100,y=100,legend = c("Ipswich","Rowley 1","Rowley 2"),pch = 16,col=unique(model$Pop),xpd = .25)
```
  
### Cumulative Variance Plot  
```{r}
out <- prcomp(t(gc))
vars <- apply(out$x, 2, var)  
props <- vars / sum(vars)
cumsum(props)
plot(cumsum(props)~c(1:24))
```

### TWO STEP DAPC: first create discriminant function from TP 9 samples and predict coordinates on df for day 80 samples.  
  
**Creating DF by treatment with first timepoint**  
```{r}
early_time_counts <- gc[,model$Day == 9]
early_time_meta <- model[model$Day == 9,]

dapc_treatment_10<-dapc(t(early_time_counts),early_time_meta$treatment,n.pca=9,n.da=2)
# PCs = 8
# clusters = 1
early_time_meta$dt <- unlist(dapc_treatment_10$ind.coord[,1])

ggplot(early_time_meta,aes(dt,fill=as.factor(treatment),colour=as.factor(treatment))) + 
  geom_density(alpha=0.1) + xlim(-8,8) + 
  labs(title="Discriminant Function for Treatment on Day 9 (based on 8 PCs)",
       x="Discriminant function 1",
       colour="Treatment",
       fill="Treatment") +
  theme_bw() +
  scale_color_manual(values=c("deepskyblue2","firebrick1")) +
  scale_fill_manual(values=c("deepskyblue2","firebrick1"))
```

Looking at which genes are driving the patterns between the two treatments
```{r}
contrib_treatment <- loadingplot(dapc_treatment_10$var.contr, axis=1,thres=.05, lab.jitter=1)
```

Looking at the most important locus based on loading
```{r}
head(gc)
gc_majorLoading <- gc[row.names(gc) == "LOC111102518",]
majorLoading_counts <- as.data.frame(cbind(Trt=model$Treatment,Time=model$Time,SFV=model$SFV,Pop=model$Pop,(LOC111104151=gc_majorLoading)))

ggplot(majorLoading_counts,aes(x=as.factor(SFV),y=log10(LOC111104151))) + geom_boxplot() + 
  labs(x= c("SFV"))
```

**Mapping Day 80 samples**  
```{r}
late_time_counts <- gc[,model$Day == 80]
late_time_meta <- model[model$Day == 80,]

predict_values <- predict.dapc(dapc_treatment_10,t(late_time_counts))
late_time_meta$dt <-unlist(predict_values$ind.scores[,1])

whole_meta<- rbind(early_time_meta,late_time_meta)

ggplot(whole_meta,aes(dt,fill=as.factor(interaction(Day,treatment)),colour=as.factor(interaction(Day,treatment)))) + 
  geom_density(alpha=0.1) + xlim(-8,8) + 
  labs(title="Discriminant Function for Treatment on Day 9 - Mapped with Day 80 Samples",
       x="Discriminant function 1",
       colour="Day.Treatment",
       fill="Day.Treatment") +
  theme_bw() +
  scale_color_manual(values=c("deepskyblue2","blue4","firebrick1","darkred")) +
  scale_fill_manual(values=c("deepskyblue2","blue4","firebrick1","darkred"))

```

### DAPC based on combined time*treatment factor  
  
**Creating DF by treatment with first timepoint**  
```{r}
dapc_SFV_10<-dapc(t(gc),model$SFV,n.pca=8,n.da=3)
# PCs = 10
# clusters = 3
output <- data.frame(Trt=model$Treatment,Time=model$Time,dapc_SFV_10$ind.coord)

ggplot(output,aes(x=LD1,y=LD2,fill=as.factor(interaction(Trt,Time)),colour=as.factor(interaction(Trt,Time)))) + 
  geom_point(aes(size=5)) + #geom_density(alpha=0.1) + #xlim(-28,28) + 
  labs(title="DAPC for Treatment*Time Combination",
       x="Discriminant function 1",
       y="Discriminant function 2",
       colour="Treatment",
       fill="Treatment") +
    theme_bw() +
  scale_color_manual(values=c("deepskyblue2","blue4","firebrick1","darkred")) +
  scale_fill_manual(values=c("deepskyblue2","blue4","firebrick1","darkred"))
```

Looking at contribution of individual genes
```{r}
contrib_SFV <- loadingplot(dapc_SFV_10$var.contr, axis=2,thres=.07, lab.jitter=1)
```
