---
title: "AE17 EPF pH Timeseries Analysis"
author: "Alan Downey-Wall"
date: "Aug 11, 2019"
output: 
    github_document
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(mgcv)
library(dplyr)
library(car)
library(lme4)
library(lmerTest)
library(kableExtra)
library(multcomp)
library(multcompView)
library(factoextra)
source("/home/downeyam/R/basicR_functions.R")
```

## Data Input and Formatting

**RData version of data**
```{r}
pheno <- readRDS("/home/downeyam/Github/2017OAExp_Oysters/input_files/Phenotype/Exposure_SummaryPhenotypeData_without81.RData")
pheno2 <- readRDS("/home/downeyam/Github/2017OAExp_Oysters/input_files/Phenotype/finalFullPhenotypeData.RData")
pheno_red <- pheno[pheno$timepoint != 81,]
```

**Creating subtable for sample info for BCO-DMO Database**
```{r}
colnames(pheno)
pheno$pCO2_cat <- as.character(pheno$pCO2.x)
pheno$pCO2_cat[pheno$pCO2_cat == "2800"] <- "High_OA"
## Subsetting full pheno for BCO-DMO data
col <- c("ID","sample_date","timepoint","pCO2_cat","shelf","tank","WaterSample_ID","PopOrigin","sequenced","width","length","wet_wgtcorr","dry_wgt","Len_Wgt_Index","HN_Index")
pheno_ind <- subset(pheno,select =col)
colnames(pheno_ind)[4] <- "treatment"
colnames(pheno_ind)[7] <- "tankID"
colnames(pheno_ind)[8] <- "collection_site"
write.csv(pheno_ind,"/home/downeyam/Github/2017OAExp_Oysters/input_files/Phenotype/AE17_SampleInfo.csv")
```

  
## Plans
  
* Look at EPF pH for all timepoints and treatments (6 tp, 3 treatments, 108 individuals, mostly balanced)  
* Look at EPF pH for all individuals we sequenced (2 tp, 2 treatments, 24 individuals)  

## Analysis of EPF pH  - Total data  
  
* Filtering out NAs (one entry) and including only timepoints from the exposure (not acclimation)
```{r echo=FALSE}
epf_exp <- pheno_red[!is.na(pheno_red$EPF_pH),]
epf_exp <-epf_exp[as.numeric(epf_exp$timepoint) > 0,]
```

**Exploratory plots**
```{r echo = FALSE}
n_sample <- epf_exp %>% group_by(pCO2_fac,timepoint_fac) %>% summarise(n=length(pCO2_fac),y.pos=quantile(pH_meas)[4])

give.n <- function(x){
   return(c(y = 7.8, label = length(x)))
}
trt_color <- c("deepskyblue3", "darkorange", "firebrick")
ggplot(epf_exp,aes(y=EPF_pH,x=timepoint_fac,group=interaction(timepoint_fac,pCO2_fac),colour=pCO2_fac)) + 
  geom_boxplot() +
  scale_color_manual(values=trt_color) + 
  theme_bw() +
  stat_summary(fun.data = give.n,
               geom = "text",
               position = position_dodge(width = 0.9), vjust = -0.25) + 
  labs(title="EPF pH",x="Time",y="EPF pH",colour="pCO2")

#Plotting some potential random effects
# Site
ggplot(epf_exp,aes(y=EPF_pH,x=PopOrigin)) + geom_boxplot()
# Some site by site variability, but nother looks significant
# Shelf
ggplot(epf_exp,aes(y=EPF_pH,x=as.factor(shelf))) + geom_boxplot()
# Maybe a shelf effect, but this is expected since it's nested in treatment
#Tank*Shelf
ggplot(epf_exp,aes(y=EPF_pH,x=interaction(as.factor(shelf),as.factor(tank)))) + geom_boxplot()
#Very noisy by tank, and it is a a little difficult to pull out which tanks are which treatments. However, looking at shelf 4 v. 6  (both 2800 treatment), there seems to be a definite shelf and maybe tank effect.
```
  
### **Statistical Analysis**  
  
#### **EPF pH total data - treatment (factor) time (continuous)**  
**Overview**  
Test: ANCOVA - Full Model (Fixed and Random Factors)  
* Predictors (fixed effects): Treatment (pCO2_fac) and Time (timepoint)  
* Random Factors: Population, Shelf, Tank (nested in Shelf)   
* Tested for normality and variance assumptions  *

**Full Model**  
```{r}
epfAllTP_full <- lmer(EPF_pH~pCO2_fac*timepoint + (1|PopOrigin) + (1|shelf/tank),data=epf_exp) 
# Warning singular fit - Indicates model is overfit and some of the the random effects will need to be removed. I use a ranova below to determine which effects are uninformative and are candidates for removal.

# Check distribution of residuals
plot(epfAllTP_full)
# They look fairly homoscedastic
qqnorm(resid(epfAllTP_full))
qqline(resid(epfAllTP_full))

anova(epfAllTP_full)
# Nothing significant here (although the interaction is close)
ranova(epfAllTP_full)
# We should remove population and shelf as random factors and rerun analysis
```
  
**Rerun Model without Population and Shelf**
```{r}
epfAllTP_red <- lmer(EPF_pH~pCO2_fac*timepoint + (1|tank:shelf),data=epf_exp)
# Prevented singular fit issue this time

# Check distribution of residuals
plot(epfAllTP_red)
# They look fairly homoscedastic
qqnorm(resid(epfAllTP_red))
qqline(resid(epfAllTP_red))
# Fairly normal
# Check autocorrelation
ggplot(epf_exp,aes(y=resid(epfAllTP_red),x=timepoint_fac,group=interaction(pCO2_fac,timepoint_fac),colour=pCO2_fac)) + geom_boxplot()
ggplot(epf_exp,aes(y=resid(epfAllTP_red),x=timepoint,colour=pCO2_fac)) + geom_point()
# Nothing obvious

# Compare full model to one with tank as the only random effect
anova(epfAllTP_full,epfAllTP_red)
ranova(epfAllTP_red) # Significant tank effect
# Comparing the two models  (full and w/ tank only there is no difference)
# We keep the simpler model

summary(epfAllTP_red) # Significant tp:2800 interaction, although not very satisfying if we want to look within tps
```
  
#### **EPF pH total data - treatment (continuous) and time (continuous)**  
**Overview**  
Test: Regression
* Predictors: Mean Tank Treatment (pCO2_calc) and Time (timepoint)  
* Random Factors: Population, Shelf, Tank  
* Tested for normality and variance assumptions  
  
```{r}
# Had to scale these values due to large differences in variance
whole_cont_full <- lmer(EPF_pH~scale(epf_exp$pCO2_calc)*scale(epf_exp$timepoint) + (1|PopOrigin) + (1|tank:shelf) ,data=epf_exp)
# Warning -  looks like this is driven by too many explanatory variables
ranova(whole_cont_full)
# Looks like we can remove population
# Ill start by removing population and shelf and reruning
```

**Rerun model with just tank as random factor**
```{r}
whole_cont_tank <- lmer(EPF_pH~scale(epf_exp$pCO2_calc)*scale(epf_exp$timepoint) + (1|tank:shelf) ,data=epf_exp)
whole_cont_tank <- lmer(EPF_pH~scale(epf_exp$pCO2_calc)*scale(epf_exp$timepoint) + (1|shelf) ,data=epf_exp)
# Better no warning this time 

# Check distribution of residuals
plot(whole_cont_tank)
# They look fairly homoscedastic
qqnorm(resid(whole_cont_tank))
qqline(resid(whole_cont_tank))
# Fairly normal

summary(whole_cont_tank)
# Treatment and intereaction significant.
```
  
#### **EPF pH total data - treatment (factor) time (factor)**  
**Overview**  
Test: 2-way ANOVA - Full Model (Fixed and Random Factors)  
* Prdictors (fixed effects): Treatment (pCO2_fac) and Time (timepoint)  
* Random Factors: Population, Shelf, Tank (nested in Shelf)   
* Tested for normality and variance assumptions  *

**Full Model**  
```{r}
epfAllTP_full_1 <- lmer(EPF_pH~pCO2_fac*timepoint_fac + (1|PopOrigin) + (1|shelf/tank),data=epf_exp) 
epfAllTP_full_2 <- lmer(EPF_pH~pCO2_fac*timepoint_fac + (1|PopOrigin) + (1|shelf/tankID),data=epf_exp)
epfAllTP_full_3 <- lmer(EPF_pH~pCO2_fac*timepoint_fac + (1|PopOrigin) + (1|shelf),data=epf_exp)
epfAllTP_full_4 <- lmer(EPF_pH~pCO2_fac*timepoint_fac + (1|PopOrigin) + (1|tank:shelf),data=epf_exp)
epfAllTP_full_5 <- lmer(EPF_pH~pCO2_fac*timepoint_fac + (1|PopOrigin) + (1|tankID),data=epf_exp)

# Model 1 (Ideal Model) : too complex leads to singularity of random effects
# Model 1 and Model are equivalent and lead to singularity issues
# Model 4 and 5 equivalent
# Start with model 5

summary(epfAllTP_full)
epfAllTP_full <- lmer(EPF_pH~pCO2_fac*timepoint_fac + (1|PopOrigin) + (1|tank/shelf),data=epf_exp) 
step(epfAllTP_full)
# Warning singular fit - Indicates model is overfit and some of the the random effects will need to be removed. I use a ranova below to determine which effects are uninformative and are candidates for removal.

# Check distribution of residuals
plot(epfAllTP_full)
# They look fairly homoscedastic
qqnorm(resid(epfAllTP_full))
qqline(resid(epfAllTP_full))

anova(epfAllTP_full)
# Interaction is significant here 
ranova(epfAllTP_full_5)
# We should remove population and shelf as random factors and rerun analysis
```
  
**Rerun Model without Population and Shelf**
```{r}
epfAllTP_red_simple <- lm(EPF_pH~pCO2_fac*timepoint_fac,data=epf_exp)
plot(epfAllTP_red_simple)
epf_exp$tankID <- as.factor(interaction(epf_exp$tank,epf_exp$shelf))

boxplot(resid(epfAllTP_red_simple)~epf_exp$tankID)
boxplot(resid(epfAllTP_red_simple)~epf_exp$PopOrigin)
boxplot(resid(epfAllTP_red_simple)~epf_exp$shelf)

epfAllTP_red <- lmer(EPF_pH~pCO2_fac*timepoint_fac + (1|tank:shelf),data=epf_exp)
# Prevented singular fit issue this time
anova(epfAllTP_red)
  # Check distribution of residuals
plot(epfAllTP_red)
# They look fairly homoscedastic
qqnorm(resid(epfAllTP_red))
qqline(resid(epfAllTP_red))
# Fairly normal
# Nothing obvious
anova(epfAllTP_red)
#Anova(epfAllTP_red,type = c("3"))
#I think this should be reasonably well balanced, so we don't need to run type 3 ANOVA

# Compare full model to one with tank as the only random effect
anova(epfAllTP_full,epfAllTP_red)
ranova(epfAllTP_red) # Significant tank effect
# Comparing the two models  (full and w/ tank only there is no difference)
# We keep the simpler model


## Planned comparisons ##
group <- paste0(pheno_red$pCO2_fac,":",pheno_red$timepoint_fac)
mod_matrix <- model.matrix(epfAllTP_red)
agg_mod_matrix <- aggregate(mod_matrix~group,FUN=mean)
rownames(agg_mod_matrix) <- agg_mod_matrix$group
agg_mod_matrix <- agg_mod_matrix[,-1]

lc2 <- as.matrix(agg_mod_matrix)
#summary(glht(epfAllTP_red,linfct=lc2),adjusted(type = "fdr"))
# Version two of planned comparisons matrix (OA trtments vs ambient at all tps)

k <- rbind("400v900_1"=lc2["900:1",]-lc2["400:1",],
           "400v2800_1"=lc2["2800:1",]-lc2["400:1",],
           "400v900_2"=lc2["900:2",]-lc2["400:2",],
           "400v2800_2"=lc2["2800:2",]-lc2["400:2",],
           "400v900_9"=lc2["900:9",]-lc2["400:9",],
           "400v2800_9"=lc2["2800:9",]-lc2["400:9",],
           "400v900_22"=lc2["900:22",]-lc2["400:22",],
           "400v2800_22"=lc2["2800:22",]-lc2["400:22",],
           "400v900_50"=lc2["900:50",]-lc2["400:50",],
           "400v2800_50"=lc2["2800:50",]-lc2["400:50",],
           "400v900_79"=lc2["900:79",]-lc2["400:79",],
           "400v2800_79"=lc2["2800:79",]-lc2["400:79",]
           )
summary(glht(epfAllTP_red,linfct=k),adjusted(type = "fdr"))

# A few significant here
```
  
#### **Relative EPF pH total data - treatment (factor) time (factor)**  
**Overview**  
Test: 2-way ANOVA - Full Model (Fixed and Random Factors)  
* Prdictors (fixed effects): Treatment (pCO2_fac) and Time (timepoint)  
* Random Factors: Population, Shelf, Tank (nested in Shelf)   
* Tested for normality and variance assumptions  *

**Full Model**  
```{r}
ggplot(epf_exp,aes(y=EPF_envAdj,x=timepoint_fac,group=interaction(timepoint_fac,pCO2_fac),colour=pCO2_fac)) + geom_boxplot()
epfAllTP_full <- lmer(EPF_envAdj~pCO2_fac*timepoint_fac + (1|PopOrigin) + (1|shelf/tank),data=epf_exp) 
# Warning singular fit - Indicates model is overfit and some of the the random effects will need to be removed. I use a ranova below to determine which effects are uninformative and are candidates for removal.

# Check distribution of residuals
plot(epfAllTP_full)
# They look fairly homoscedastic
qqnorm(resid(epfAllTP_full))
qqline(resid(epfAllTP_full))

anova(epfAllTP_full)
# Interaction is significant here 
ranova(epfAllTP_full)
# We should remove population and shelf as random factors and rerun analysis
```
  
**Rerun Model without Population and Shelf**
```{r}
epfAllTP_red <- lmer(EPF_envAdj~pCO2_fac*timepoint_fac + (1|tank:shelf),data=epf_exp)
# Prevented singular fit issue this time
anova(epfAllTP_red)
# Check distribution of residuals
plot(epfAllTP_red)
# They look fairly homoscedastic
qqnorm(resid(epfAllTP_red))
qqline(resid(epfAllTP_red))
# Fairly normal
# Nothing obvious
anova(epfAllTP_red)
#Anova(epfAllTP_red,type = c("3"))
#I think this should be reasonably well balanced, so we don't need to run type 3 ANOVA

# Compare full model to one with tank as the only random effect
anova(epfAllTP_full,epfAllTP_red)
ranova(epfAllTP_red) # Significant tank effect
# Comparing the two models  (full and w/ tank only there is no difference)
# We keep the simpler model


## Planned comparisons ##
group <- paste0(pheno_red$pCO2_fac,":",pheno_red$timepoint_fac)
mod_matrix <- model.matrix(epfAllTP_red)
agg_mod_matrix <- aggregate(mod_matrix~group,FUN=mean)
rownames(agg_mod_matrix) <- agg_mod_matrix$group
agg_mod_matrix <- agg_mod_matrix[,-1]

lc2 <- as.matrix(agg_mod_matrix)
summary(glht(epfAllTP_red,linfct=lc2),adjusted(type = "fdr"))
# Version two of planned comparisons matrix (OA trtments vs ambient at all tps)

k <- rbind("400v900_1"=lc2["900:1",]-lc2["400:1",],
           "400v2800_1"=lc2["2800:1",]-lc2["400:1",],
           "400v900_2"=lc2["900:2",]-lc2["400:2",],
           "400v2800_2"=lc2["2800:2",]-lc2["400:2",],
           "400v900_9"=lc2["900:9",]-lc2["400:9",],
           "400v2800_9"=lc2["2800:9",]-lc2["400:9",],
           "400v900_22"=lc2["900:22",]-lc2["400:22",],
           "400v2800_22"=lc2["2800:22",]-lc2["400:22",],
           "400v900_50"=lc2["900:50",]-lc2["400:50",],
           "400v2800_50"=lc2["2800:50",]-lc2["400:50",],
           "400v900_79"=lc2["900:79",]-lc2["400:79",],
           "400v2800_79"=lc2["2800:79",]-lc2["400:79",]
           )
summary(glht(epfAllTP_red,linfct=k),adjusted(type = "fdr"))

# A few significant here
```
  
## Figure

```{r Map with all treatment levels, echo=FALSE, fig.height=10}
c_mean <- mean(epf_exp$Tank_pH[epf_exp$pCO2_fac==400])
oa_900_mean <- mean(epf_exp$Tank_pH[epf_exp$pCO2_fac==900])
oa_2800_mean <- mean(epf_exp$Tank_pH[epf_exp$pCO2_fac==2800])
par(mfrow=c(2,1))
#### Measured EPF ph plot ####
treatSeq_means <- aggregate(EPF_pH~timepoint+pCO2_fac,epf_exp,FUN=mean)
treat_means_ctrl <- treatSeq_means[treatSeq_means$pCO2_fac == "400",]
treat_means_ctrl$timepoint[2] <- 4
treat_means_oa_900 <- treatSeq_means[treatSeq_means$pCO2_fac == "900",]
treat_means_oa_900$timepoint <- treat_means_oa_900$timepoint + 1
treat_means_oa_900$timepoint[2] <- 5
treat_means_oa_2800 <- treatSeq_means[treatSeq_means$pCO2_fac == "2800",]
treat_means_oa_2800$timepoint <- treat_means_oa_2800$timepoint + 2
treat_means_oa_2800$timepoint[2] <- 6

treatSeq_SE <- aggregate(EPF_pH~timepoint+pCO2_fac,epf_exp,FUN=se)
#treatSeq_SE[,3] <- treatSeq_SE[,3]*1.96
treat_SE_ctrl <- treatSeq_SE[treatSeq_SE$pCO2_fac == "400",]
treat_SE_oa_900 <- treatSeq_SE[treatSeq_SE$pCO2_fac == "900",]
treat_SE_oa_900$timepoint <- treat_SE_oa_900$timepoint + 2
treat_SE_oa_2800 <- treatSeq_SE[treatSeq_SE$pCO2_fac == "2800",]
treat_SE_oa_900$timepoint <- treat_SE_oa_900$timepoint + 4

# Base plot with 400
bp <- plot(treat_means_ctrl$EPF_pH~as.numeric(as.character(treat_means_ctrl$timepoint)),ylab="EPF pH (NBS)",xlab="Time (Days)",
           col="lightblue4",pch=16,cex=1.5,
           ylim = c(6.5,8.1),xlim=c(0,82),bty="n")

# Mean environment lines underneath other lines    
abline(h=c_mean,col="lightblue4",lty=2) # control
abline(h=oa_900_mean,col="goldenrod",lty=2) # 900
abline(h=oa_2800_mean,col="tomato",lty=2) # 2800

# 400 EPF lines
lines(treat_means_ctrl$EPF_pH~as.numeric(as.character(treat_means_ctrl$timepoint)),
      col="lightblue4")
arrows(x0 = as.numeric(as.character(treat_means_ctrl$timepoint)),
       x1 = as.numeric(as.character(treat_means_ctrl$timepoint)),
       y0 = c(treat_means_ctrl$EPF_pH - treat_SE_ctrl$EPF_pH),
       y1 = treat_means_ctrl$EPF_pH + treat_SE_ctrl$EPF_pH,
       angle = 90, len = 0.05,
       code = 3, xpd = NA, lwd = 2,
       col="lightblue4")

# 900 EPF lines
points(treat_means_oa_900$EPF_pH~as.numeric(as.character(treat_means_oa_900$timepoint)),
       col="goldenrod",pch=15,cex=1.2)
lines(treat_means_oa_900$EPF_pH~as.numeric(as.character(treat_means_oa_900$timepoint)),
      col="goldenrod")
arrows(x0 = as.numeric(as.character(treat_means_oa_900$timepoint)),
       x1 = as.numeric(as.character(treat_means_oa_900$timepoint)),
       y0 = c(treat_means_oa_900$EPF_pH - treat_SE_ctrl$EPF_pH),
       y1 = treat_means_oa_900$EPF_pH + treat_SE_ctrl$EPF_pH,
       angle = 90, len = 0.05,
       code = 3, xpd = NA, lwd = 2,
       col="goldenrod")

# 2800 treatment 
points(treat_means_oa_2800$EPF_pH~as.numeric(as.character(treat_means_oa_2800$timepoint)),
       col="tomato",pch=17,cex=1.2)
lines(treat_means_oa_2800$EPF_pH~as.numeric(as.character(treat_means_oa_2800$timepoint)),
      col="tomato")
arrows(x0 = as.numeric(as.character(treat_means_oa_2800$timepoint)),
       x1 = as.numeric(as.character(treat_means_oa_2800$timepoint)),
       y0 = c(treat_means_oa_2800$EPF_pH - treat_SE_oa_2800$EPF_pH),
       y1 = treat_means_oa_2800$EPF_pH + treat_SE_oa_2800$EPF_pH,
       angle = 90, len = 0.05, code = 3, xpd = NA, lwd = 2,
       col="tomato")

# Legend
legend(x=2,y=6.9,
       legend = c("Ambient","OA 900","OA 2800"),
       col = c("lightblue4","goldenrod","tomato"),
       pch=c(16,15,17),
       cex = 1.2,
       lwd = 2,
       bty = "n")
text(x=-8,y=8.37,label="A",cex = 2, xpd = NA)

text(x=5,y=8,label="*",cex = 2,col="tomato", xpd = NA)
text(x=51,y=8,label="*",cex = 2,col="tomato", xpd = NA)
text(x=80,y=8,label="*",cex = 2,col="tomato", xpd = NA)

#### EPF - Environment plot ####
treatSeq_means <- aggregate(EPF_envAdj~timepoint+pCO2_fac,epf_exp,FUN=mean)
treat_means_ctrl <- treatSeq_means[treatSeq_means$pCO2_fac == "400",]
treat_means_ctrl$timepoint[2] <- 4
treat_means_oa_900 <- treatSeq_means[treatSeq_means$pCO2_fac == "900",]
treat_means_oa_900$timepoint <- treat_means_oa_900$timepoint + 1
treat_means_oa_900$timepoint[2] <- 5

treat_means_oa_2800 <- treatSeq_means[treatSeq_means$pCO2_fac == "2800",]
treat_means_oa_2800$timepoint <- treat_means_oa_2800$timepoint + 2
treat_means_oa_2800$timepoint[2] <- 6
#treatSeq_SE <- aggregate(EPF_envAdj~timepoint_fac+pCO2_fac,epf_exp,FUN=sd)
treatSeq_SE <- aggregate(EPF_envAdj~timepoint+pCO2_fac,epf_exp,FUN=se)
treat_SE_ctrl <- treatSeq_SE[treatSeq_SE$pCO2_fac == "400",]
treat_SE_oa_900 <- treatSeq_SE[treatSeq_SE$pCO2_fac == "900",]
treat_SE_oa_900$timepoint <- treat_SE_oa_900$timepoint + 2
treat_SE_oa_2800 <- treatSeq_SE[treatSeq_SE$pCO2_fac == "2800",]
treat_SE_oa_900$timepoint <- treat_SE_oa_900$timepoint + 4

# Base plot with 400
bp <- plot(treat_means_ctrl$EPF_envAdj~as.numeric(as.character(treat_means_ctrl$timepoint)),
           ylab="EPF pH - Environment pH (NBS)",
           xlab="Time (Days)",pch=16,
           col="lightblue4",cex=1.2,
           ylim = c(-1,1),xlim=c(0,82),bty="n")

# Mean environment lines underneath other lines    
abline(h=0,col="black",lty=2) # control

# 400 EPF lines
lines(treat_means_ctrl$EPF_envAdj~as.numeric(as.character(treat_means_ctrl$timepoint)),
      col="lightblue4")
arrows(x0 = as.numeric(as.character(treat_means_ctrl$timepoint)),
       x1 = as.numeric(as.character(treat_means_ctrl$timepoint)),
       y0 = c(treat_means_ctrl$EPF_envAdj - treat_SE_ctrl$EPF_envAdj),
       y1 = treat_means_ctrl$EPF_envAdj + treat_SE_ctrl$EPF_envAdj,
       angle = 90, len = 0.05,
       code = 3, xpd = NA, lwd = 2,
       col="lightblue4")

# 900 EPF lines
points(treat_means_oa_900$EPF_envAdj~as.numeric(as.character(treat_means_oa_900$timepoint)),
       col="goldenrod",pch=15,cex=1.2)
lines(treat_means_oa_900$EPF_envAdj~as.numeric(as.character(treat_means_oa_900$timepoint)),
      col="goldenrod")
arrows(x0 = as.numeric(as.character(treat_means_oa_900$timepoint)),
       x1 = as.numeric(as.character(treat_means_oa_900$timepoint)),
       y0 = c(treat_means_oa_900$EPF_envAdj - treat_SE_ctrl$EPF_envAdj),
       y1 = treat_means_oa_900$EPF_envAdj + treat_SE_ctrl$EPF_envAdj,
       angle = 90, len = 0.05,
       code = 3, xpd = NA, lwd = 2,
       col="goldenrod")

# 2800 treatment 
points(treat_means_oa_2800$EPF_envAdj~as.numeric(as.character(treat_means_oa_2800$timepoint)),
       col="tomato",pch=17,cex=1.2)
lines(treat_means_oa_2800$EPF_envAdj~as.numeric(as.character(treat_means_oa_2800$timepoint)),
      col="tomato")
arrows(x0 = as.numeric(as.character(treat_means_oa_2800$timepoint)),
       x1 = as.numeric(as.character(treat_means_oa_2800$timepoint)),
       y0 = c(treat_means_oa_2800$EPF_envAdj - treat_SE_oa_2800$EPF_envAdj),
       y1 = treat_means_oa_2800$EPF_envAdj + treat_SE_oa_2800$EPF_envAdj,
       angle = 90, len = 0.05, code = 3, xpd = NA, lwd = 2,
       col="tomato")

# Legend
legend(x=65,y=0.8,
       legend = c("Ambient","OA 900","OA 2800"),
       col = c("lightblue4","goldenrod1","tomato"),
       pch=c(16,15,17),
       cex = 1.2,
       lwd = 2,
       bty = "n")
text(x=-8,y=1.3,label="B",cex = 2, xpd = NA)

text(x=2,
     y=0.95,label="*",
     cex = 2,col="lightblue4", xpd = NA)
text(x=5,
     y=0.95,label="+",
     cex = 2,col="lightblue4", xpd = NA)
text(x=10,
     y=0.95,label="***",
     cex = 2,col="lightblue4", xpd = NA)
text(x=23,
     y=0.95,label="***",
     cex = 2,col="lightblue4", xpd = NA)
text(x=50,
     y=0.95,label="*",
     cex = 2,col="lightblue4", xpd = NA)
text(x=23,
     y=0.85,label="+",
     cex = 2,col="goldenrod1", xpd = NA)
text(x=10,
     y=0.75,label="***",
     cex = 2,col="tomato", xpd = NA)
```
  
## Analysis of EPF pH -  Only samples we sequenced  

#### **Measured EPF pH total data - time as factor**  
**Overview**  
Test: 2-way ANOVA - Full Model (Fixed and Random Factors)  
* Explanatory Factors: Treatment (pCO2_fac) and Time (timepoint_fac)  
* Random Factors: Population, Shelf, Tank (nested in Shelf)   
* Tested for normality and variance assumptions  

```{r}
pheno_seq <- pheno[pheno$sequenced == "Y",]
pheno_seq$timepoint[pheno_seq$timepoint == 81] <- 80
pheno_seq$timepoint[pheno_seq$timepoint == 79] <- 80
pheno_seq$timepoint_fac <- as.factor(pheno_seq$timepoint)

seq_full <- lmer(EPF_pH~timepoint_fac*pCO2_fac + (1|PopOrigin) + (1|shelf/tank),data=pheno_seq)
# Singular fit warning

plot(seq_full)
# Some heteroscedascity but not too bad
qqnorm(resid(seq_full))
qqline(resid(seq_full))

ranova(seq_full) # None of these seem important

anova(seq_full) # Everything except pCO2 significant

# Decide to remove random factors since they aren't informative
```

**Rerun with standard lm**
```{r}
seq_fixed <- lm(EPF_pH~timepoint_fac*pCO2_fac, data=pheno_seq)

plot(seq_fixed)
# We can see more clearly that some of the mild issues with the assumptions are being driven by 1 (61).

anova(seq_fixed) # without random factors, pCO2 is now even less significant
```
  
**Rerun but remove the one problematic points**
```{r}
# New dataset w/o the one point
pheno_seq2 <- pheno_seq[row.names(pheno_seq) != c(48),]
seq_fixed_outlierRM <- lm(EPF_pH~timepoint_fac*pCO2_fac, data=pheno_seq2)
plot(seq_fixed_outlierRM)
anova(seq_fixed_outlierRM)
# On a whole this looks like a slight improvement
pheno_seq3 <- pheno_seq2[row.names(pheno_seq2) != c(3),]
seq_fixed_outlierRM <- lm(EPF_pH~timepoint_fac*pCO2_fac, data=pheno_seq3)
plot(seq_fixed_outlierRM)
anova(seq_fixed_outlierRM)
```
  
**Pairwise comparisons of fixed factor model (with points rm)**  
```{r}
seq_pairwise <- TukeyHSD(aov(EPF_pH~timepoint_fac*pCO2_fac, data=pheno_seq3))
plot(seq_pairwise)
```

#### **Relative EPF pH (EPF pH - Env. pH) total data - time as factor**  
**Overview**  
Test: 2-way ANOVA - Full Model (Fixed and Random Factors)  
* Explanatory Factors: Treatment (pCO2_fac) and Time (timepoint_fac)  
* Random Factors: Population, Shelf, Tank (nested in Shelf)   
* Tested for normality and variance assumptions  

```{r}
pheno_seq <- pheno[pheno$sequenced == "Y",]
pheno_seq$timepoint[pheno_seq$timepoint == 81] <- 80
pheno_seq$timepoint[pheno_seq$timepoint == 79] <- 80
pheno_seq$timepoint_fac <- as.factor(pheno_seq$timepoint)

seq_full <- lmer(EPF_envAdj~timepoint_fac*pCO2_fac + (1|PopOrigin) + (1|shelf/tank),data=pheno_seq)
# Singular fit warning

plot(seq_full)
# Some heteroscedascity but not too bad
qqnorm(resid(seq_full))
qqline(resid(seq_full))

ranova(seq_full) # None of these seem important

anova(seq_full) # Everything significant except time

# Decide to remove random factors since they aren't informative
```

**Rerun with standard lm**
```{r}
seq_fixed <- lm(EPF_envAdj~timepoint_fac*pCO2_fac, data=pheno_seq)

plot(seq_fixed)
# We can see more clearly that some of the mild issues with the assumptions are being driven by 1 (61).

anova(seq_fixed) # without random factors, pCO2 is now even less significant
```
  
**Rerun but remove the one problematic points**
```{r}
# New dataset w/o the one point
pheno_seq2 <- pheno_seq[row.names(pheno_seq) != c(48),]
seq_fixed_outlierRM <- lm(EPF_envAdj~timepoint_fac*pCO2_fac, data=pheno_seq2)
plot(seq_fixed_outlierRM)
anova(seq_fixed_outlierRM)
# On a whole this looks like a slight improvement
pheno_seq3 <- pheno_seq2[row.names(pheno_seq2) != c(3),]
seq_fixed_outlierRM <- lm(EPF_envAdj~timepoint_fac*pCO2_fac, data=pheno_seq3)
plot(seq_fixed_outlierRM)
anova(seq_fixed_outlierRM)
```
  
**Pairwise comparisons of fixed factor model (with points rm)**  
```{r}
seq_pairwise <- TukeyHSD(aov(EPF_envAdj~timepoint_fac*pCO2_fac, data=pheno_seq3))
plot(seq_pairwise)
```

**Final Figures**  
```{r echo=FALSE}
seq2_means <- aggregate(EPF_pH~timepoint_fac+pCO2_fac,pheno_seq3,FUN=mean)
seq2_SE <- aggregate(EPF_pH~timepoint_fac+pCO2_fac,pheno_seq2,FUN=se)
seq2_SE$ci <- seq2_SE$EPF_pH*1.96
seq2_means$comb <- interaction(seq2_means$timepoint_fac,seq2_means$pCO2_fac)
seq2_means$ci <- seq2_SE$ci

seq_labels <- multcompLetters(seq_pairwise$`timepoint_fac:pCO2_fac`[, "p adj"])$Letters
seq_labels <- seq_labels[c(1,2,4,3)]


ggplot(seq2_means,aes(y=EPF_pH,x=comb,colour=comb)) + 
  geom_point(size=5,aes(shape=pCO2_fac)) + ylim(6.5,8) +
  scale_colour_manual(values=c("lightblue4","royalblue3","tomato","darkred")) +
  geom_errorbar(aes(ymin=(EPF_pH-ci),ymax=(EPF_pH+ci)),width=.2) + 
  scale_x_discrete(limits=c("9.400","9.2800","80.400","80.2800")) + 
  labs(x="Treatment and Time",y="EPF pH (NBS)",size=6) + theme_bw() +
  theme(strip.text.x = element_text(size=15),
        legend.position="none",
        panel.grid.major.x = element_blank() ,
        strip.background = element_rect(colour=NULL, fill="white"))

# Relative EPF pH
seq2_means <- aggregate(EPF_envAdj~timepoint_fac+pCO2_fac,pheno_seq3,FUN=mean)
seq2_SE <- aggregate(EPF_envAdj~timepoint_fac+pCO2_fac,pheno_seq2,FUN=se)
seq2_SE$ci <- seq2_SE$EPF_envAdj#*1.96
seq2_means$comb <- interaction(seq2_means$timepoint_fac,seq2_means$pCO2_fac)
seq2_means$ci <- seq2_SE$ci
seq2_means$x <- c(1,4,2,5)
#saveRDS(seq2_means,"/home/downeyam/Github/2017OAExp_Oysters/input_files/Phenotype/RelativeEPFpH_sampleSubset_SummaryTable.RData")

ggplot(seq2_means,aes(y=EPF_envAdj,x=comb,fill=comb)) + 
  geom_col() + #ylim(6.5,8) +
  scale_fill_manual(values=c("lightblue4","royalblue3","tomato","darkred")) +
  geom_errorbar(aes(ymin=(EPF_envAdj-ci),ymax=(EPF_envAdj+ci)),width=.2) + 
  scale_x_discrete(limits=c("9.400","9.2800","80.400","80.2800")) + 
  labs(x="Treatment and Time",y="EPF pH (NBS)",size=6) + theme_bw() +
  theme(strip.text.x = element_text(size=15),
        legend.position="none",
        panel.grid.major.x = element_blank() ,
        strip.background = element_rect(colour=NULL, fill="white"))

```
  
Final bar plot has SE bars and significance levels based on pairwise comparison test (tukey method) based on model selected above.

## Final Summary  
  
**General Observations**: The data was analyzed primarily using ANOVAs based on linear mixed models that included tank as a random effect and pCO2 treatment and time as fixed effects. Time in this data set was primarily treated as a factor (along with pCO2 treatment) given the non linear relationship with EPF pH and time and the lack of a truly robust timeseries for most comparisons (however it might still be a good idead to examine the complete EPF pH timeseries with additional analyses). Generally, pCO2 was found to have a significant effect on EPF pH, and the interaction of time and treatment was found to also often be significant. Importantly, when just considering the two timepoints with sequence data (Day 9 and day 80) treatment, time, and the interaction were all significant.

**Break down of the final models by organizational level**  
  
Full Dataset (All timepoints, 3 treatments)  
* Best Model : EPF_pH ~ Treatment * Time + (1|Tank:Shelf)  
```{r echo=FALSE}
#kable(anova(epfAllTP_red)) %>% kable_styling()
```
  
EPF pH from sequenced individuals (2 timepoints, 2 treatments, 22 samples)  

* Best Model : EPF_pH ~ Treatment * Time + (1|Tank:Shelf)  
```{r echo=FALSE}
#kable(anova(seq_fixed_outlierRM)) %>% kable_styling()
```
  

  
