---
title: "AE17 Final Carbonate Chemistry Analysis"
author: "adowneywall"
date: "8/11/2019"
output: 
    github_document
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(mgcv)
library(dplyr)
library(car)
library(lme4)
library(lmerTest)
library(kableExtra)
library(multcomp)
library(multcompView)
library(factoextra)
source("/home/downeyam/R/basicR_functions.R")
```

**RData version of data**
```{r Data read in and configuration}
epf_wDIC_red <- readRDS("/home/downeyam/Github/2017OAExp_Oysters/input_files/Phenotype/Exposure_SummaryFullCarbChemData_without81.RData")
```

## Subsetting data for only days we sequenced (09 v 80 - All treatments)

**NOTE**: Tried random effects models but found they either led toa singular fit (over fitting problem) or they were not significant and removed. Presented below are only the best models identifies via model selection (LRT) and after making sure all assumptions were met and all include simple linear model with **Treatment**, **Time**, and **Treatment:Time** as explainatory factors. Also note that both **Treatment** and **Time** were handled as factors, and a type III ANOVA was performed due to sample/unequal sample sizes among groups.

### EPF pH
```{r}
par(mfrow=c(1,1))

epf_wDIC_red <- epf_wDIC_red[epf_wDIC_red$timepoint != 81,]
epf_wDIC_red <- epf_wDIC_red[epf_wDIC_red$timepoint != 22,]
epf_wDIC_red$timepoint_fac <- as.factor(epf_wDIC_red$timepoint)


mode_pH_f <- lm(EPF_pH ~ pCO2_fac*timepoint_fac,data=epf_wDIC_red)
Anova(mode_pH_f,type=3)
# Significant interaction

(pH_seq_pairwise <- TukeyHSD(aov(EPF_pH~timepoint_fac*pCO2_fac, data=epf_wDIC_red)))
plot(pH_seq_pairwise$`timepoint_fac:pCO2_fac`)
```

### Analysis DIC 
```{r}
mod_ca <- lm(log(EPF_DIC)~pCO2_fac*timepoint_fac,data=epf_wDIC_red)
Anova(mod_ca,type=3)
#Nothing significant

(DIC_seq_pairwise <- TukeyHSD(aov(log(EPF_DIC)~timepoint_fac*pCO2_fac, data=epf_wDIC_red)))
plot(DIC_seq_pairwise$`timepoint_fac:pCO2_fac`)
```

### Analysis Ca Calcite
```{r}
mod_ca <- lm(EPF_Calcite_Calc~pCO2_fac*timepoint_fac,data=epf_wDIC_red)
Anova(mod_ca,type=3)
#Significant by treatment and interaction

(Ca_seq_pairwise <- TukeyHSD(aov(EPF_Calcite_Calc~timepoint_fac*pCO2_fac, data=epf_wDIC_red)))
plot(Ca_seq_pairwise$`timepoint_fac:pCO2_fac`)
```

## Figures for sequenced samples
```{r echo=FALSE, fig.height=10}
par(mfrow=c(3,1),mar=c(5,5.5,2,1),oma=c(3,2,2,0))

### Calcite #####
treatSeq_means <- aggregate(EPF_Calcite_Calc~pCO2_fac+timepoint,epf_wDIC_red,FUN=mean)
treatSeq_SE <- aggregate(EPF_Calcite_Calc~pCO2_fac+timepoint,epf_wDIC_red,FUN=se)
treatSeq_SE$ci <- treatSeq_SE$EPF_Calcite_Calc*1.96

seq_labels <- multcompLetters(pH_seq_pairwise$`timepoint_fac:pCO2_fac`[, "p adj"])$Letters
seq_labels <- seq_labels[c(1,2,4,3)]

bp <- barplot(treatSeq_means$EPF_Calcite_Calc,ylim=c(0,35),
              col=c("lightblue4","goldenrod","tomato"),
              ylab = "EPF Calcite",
              xpd = FALSE,cex.axis = 2,cex.lab = 1.8)
#xlab(cex=2)
abline(h=1,lty=2,cex=2)
arrows(x0 = bp, x1 = bp,
       y0 = treatSeq_means$EPF_Calcite - treatSeq_SE$ci,
       y1 = treatSeq_means$EPF_Calcite + treatSeq_SE$ci,
       angle = 90, len = 0.1, code = 3, xpd = NA,cex=3)
text(x=-.7,y=40,label="A",cex = 3, xpd = NA)
#text(x = bp, y = -2.5, label = c("Control","OA\n1000","OA\n2800","Control","OA\n1000","OA\n2800"), xpd = NA,cex=1.8)
#text(x = mean(bp[1:3]), y = -7, label = "Day 9", xpd = NA,cex=1.8)
#text(x = mean(bp[4:6]), y = -7, label = "Day 80", xpd = NA,cex=1.8)
text(x = bp, y = 40, label = c("n=2","n=3","n=3","n=1","n=3","n=3"), xpd = NA,cex=1.8)
box()
#text(x = bp, y = treatSeq_means$EPF_Calcite + treatSeq_SE$EPF_Calcite, pos = 3, lab = seq_labels, xpd = NA)

### EPF pH #####
treatSeq_means <- aggregate(EPF_pH~pCO2_fac+timepoint,epf_wDIC_red,FUN=mean)
treatSeq_SE <- aggregate(EPF_pH~pCO2_fac+timepoint,epf_wDIC_red,FUN=se)
treatSeq_SE$ci <- treatSeq_SE$EPF_pH*1.96
seq_labels <- multcompLetters(pH_seq_pairwise$`timepoint_fac:pCO2_fac`[, "p adj"])$Letters
seq_labels <- seq_labels[c(6,2,4,1,3,5)]

bp <- barplot(treatSeq_means$EPF_pH,ylim=c(6.5,8.1),
              col=c("lightblue4","goldenrod","tomato"),
              ylab = "EPF pH",
              xpd = FALSE,cex.axis = 1.8,cex.lab=1.8)
arrows(x0 = bp, x1 = bp,
       y0 = treatSeq_means$EPF_pH - treatSeq_SE$ci,
       y1 = treatSeq_means$EPF_pH + treatSeq_SE$ci,
       angle = 90, len = 0.1, code = 3, xpd = NA)
text(x=-.7,y=8.3,label="B",cex = 3, xpd = NA)
#text(x = bp, y = 6.4, label =  c("Control","OA\n1000","OA\n2800","Control","OA\n1000","OA\n2800"), xpd = NA,cex=1.8)
#text(x = mean(bp[1:3]), y = 6.2, label = "Day 9", xpd = NA,cex=1.8)
#text(x = mean(bp[4:6]), y = 6.2, label = "Day 80", xpd = NA,cex=1.8)
box()
text(x = bp, y = treatSeq_means$EPF_pH + treatSeq_SE$ci, pos = 3, lab = seq_labels, xpd = NA,cex=1.8)

### DIC ####
treatSeq_means <- aggregate(EPF_DIC~pCO2_fac+timepoint,epf_wDIC_red,FUN=mean)
treatSeq_SE <- aggregate(EPF_DIC~pCO2_fac+timepoint,epf_wDIC_red,FUN=se)
treatSeq_SE$ci <- treatSeq_SE$EPF_DIC*1.96
seq_labels <- multcompLetters(DIC_seq_pairwise$`timepoint_fac:pCO2_fac`[, "p adj"])$Letters
seq_labels <- seq_labels[c(1,2,4,3)]

bp <- barplot(treatSeq_means$EPF_DIC,ylim = c(0,50000),
              col=c("lightblue4","goldenrod","tomato"),
              ylab = "EPF DIC",
              xpd = FALSE,cex.axis = 1.8,cex.lab=1.8)
arrows(x0 = bp, x1 = bp,
       y0 = treatSeq_means$EPF_DIC - treatSeq_SE$ci,
       y1 = treatSeq_means$EPF_DIC + treatSeq_SE$ci,
       angle = 90, len = 0.1, code = 3, xpd = NA)
text(x=-.7,y=58000,label="C",cex = 3, xpd = NA)
text(x = bp, y = -8000, label = c("Control","OA\n1000","OA\n2800","Control","OA\n1000","OA\n2800"), xpd = NA,cex=1.8)
text(x = mean(bp[1:3]), y = -20000, label = "Day 9", xpd = NA,cex=1.8)
text(x = mean(bp[4:6]), y = -20000, label = "Day 80", xpd = NA,cex=1.8)
box()
#text(x = bp, y = treatSeq_means$EPF_DIC + treatSeq_SE$EPF_DIC, pos = 3, lab = seq_labels, xpd = NA)
#table(epf_wDIC_red$timepoint_fac,epf_wDIC_red$pCO2_fac)
```



