---
title: "AE17 Calcification Full Workup"
author: "adowneywall"
date: "8/21/2019"
output: github_document
editor_options:     
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(lmerTest)
library(car)
source("/home/downeyam/R/basicR_functions.R")
library(RColorBrewer)
pal <- brewer.pal(n = 12, name = 'Paired')
col_perm <- c(pal[1:2],pal[5:6],pal[12])
```

### Data
```{r}
# Final version of bouyant weights
cal<-read.csv("/home/downeyam/Github/2017OAExp_Oysters/input_files/Phenotype/AE17_BouyantWeight_Final.csv")
# Final version of sample data
samp<-read.csv("/home/downeyam/Github/2017OAExp_Oysters/input_files/Phenotype/AE17_CollectionInfo.csv")
# Final EPF fluid data 
pheno <- readRDS("/home/downeyam/Github/2017OAExp_Oysters/input_files/Phenotype/AE17_summaryPhenotype_exposure.RData")
pheno_red <- pheno[pheno$timepoint != 81,]
pheno_red <- subset(pheno_red,select=c(1,30,37))

# Merge with sample sheet to include site info
samp_red <- subset(samp,select=c(1,9,38,39,40,41,42))
cal <- subset(cal,select=c(1:14,15,17,21,23,27,29,39:51))
cal <- merge(samp_red,cal)

# Filter complete dataset
# Remove individuals that don't have values for the third bouyant weight
cal_red <- cal[!is.na(cal$bw2_bw3_daily_sizeCorr),]
# Remove the extra oysters used for the calcein experiment and from tp 81
cal_red2 <-  cal_red2[cal_red$sample_date < 20170824,]
```

**Preliminary plot and model of points**
```{r}
## Including calcein samples
# Boxplot
ggplot(cal_red,aes(x=as.factor(pCO2),y=bw2_bw3_daily,group=as.factor(pCO2))) + geom_boxplot()
# Table of points
table(cal_red$pCO2)
# ANOVA and post hoc
out.aov <- aov(cal_red$bw2_bw3_daily~as.factor(cal_red$pCO2))
summary(out.aov)
TukeyHSD(out.aov)
## Including additional samples
# Boxplot
ggplot(cal_red2,aes(x=as.factor(pCO2),y=bw2_bw3_daily,group=as.factor(pCO2))) + geom_boxplot()
# Table of points
table(cal_red2$pCO2)
# ANOVA and post hoc
out.aov <- aov(cal_red2$bw2_bw3_daily~as.factor(cal_red2$pCO2))
TukeyHSD(out.aov)
## Removing additional samples
#Boxplot (pco2 as factor)
ggplot(cal_red2,aes(x=as.factor(pCO2),y=bw2_bw3_daily,group=as.factor(pCO2))) + geom_boxplot()
# Table of points
table(cal_red2$pCO2)
# ANOVA and post hoc
out.aov <- aov(cal_red2$bw2_bw3_daily~as.factor(cal_red2$pCO2))
summary(out.aov)
TukeyHSD(out.aov)
# box plot (pco2 as as continuous variable)
ggplot(cal_red2,aes(x=pCO2,y=bw2_bw3_daily,group=as.factor(pCO2))) + geom_boxplot()

## Scaling by oyster size
# box plot (pco2 as as continuous variable)
ggplot(cal_red2,aes(x=pCO2,y=bw2_bw3_daily/dry2_final,group=as.factor(pCO2))) + geom_boxplot()
# ANOVA and post hoc
out.aov <- aov(bw2_bw3_daily/dry2_final~as.factor(pCO2),data=cal_red2)
summary(out.aov)
TukeyHSD(out.aov)
# Linear model
out.lm <- lm(bw2_bw3_daily/dry2_final~pCO2_calc,data=cal_red2)
summary(out.lm)
ggplot(cal_red2,aes(x=pCO2_calc,y=bw2_bw3_daily/dry2_final,colour=as.factor(pCO2))) + geom_point()
```

### Primary Variables 

**Fixed Effects**

* Treatment 
  * pCO2_fac = Treating pCO2 like a factor
  * pCO2_calc = Treating pCO2 as continuous variable using tank mean values

* Time
  * Timepoint_fac =  Two (or three) timepoints as factors
  * timepoint = Two timepoint as continous variable

**Random Effects**

* Pop : 3 sites -  roughly 1/3 oysters per site
* Shelf : 6 shelfs (nested in treatment) - 12-18 oysters per shelf
* Tank : 3 tanks per shelf (nested) - 2 -3 oysters per tank

### Tables

**All timepoints including calcein**
```{r}
# Table
table(cal_red$sample_date,cal_red$pCO2)
# Total number of samples
nrow(cal_red)
```

**All timepoints w/o calcein**  
Rationale explained in notes at end
```{r}
# Table
table(cal_red2$sample_date,cal_red2$pCO2)
# Total number of samples
nrow(cal_red2)
```

**Timepoints with all data (final dataset)**  
Rationale explained in notes at end
```{r}
# Table
table(cal_red2$sample_date,cal_red2$pCO2)
# Total number of samples
nrow(cal_red2)
```

## Preliminary Plots with Calcification Data
```{r}
### Plotting Calcification Data
## Calcification Adj -- with Time -- Full Data
gg <- ggplot(cal_red,aes(x=pCO2,y=bw2_bw3_daily/dry2_final,
                         group=interaction(pCO2,sample_date),
                         colour=interaction(pCO2,sample_date))) 
gg + geom_boxplot() +
  labs(colour="pCO2.Timepoint",x="pCO2",title="All timepoints")

ggplot(cal_red,aes(x=interaction(tank,shelf),y=bw2_bw3_daily/dry2_final,colour=as.factor(pCO2))) + geom_boxplot()+ labs(colour="pCO2",x="Tanks.Shelf",title="All timepoints")
```

## Calcification vs Environment and Time - Statistical Analysis

**ANOVA**
```{r}
cal_red2$calcification <- cal_red2$bw2_bw3_daily/cal_red2$dry2_final

# Full model- All points
cal_full <- lmer(calcification~as.factor(pCO2) + (1|PopOrigin) + (1|shelf/tank),data=cal_red2)
# Singular fit so we need to simplfy
anova(cal_full) # No significance
ranova(cal_full)
#None are significant / Remove pop and shelf and rerun

# Reducing model - keeping fixed factors and one random effect
cal_full_2 <- lmer(calcification~as.factor(pCO2) + (1|tank:shelf),data=cal_red2)
ranova(cal_full_2)
# Random effect still not significant but i will keep since it's close 
summary(cal_full_2)

# Fixed effects only
cal_analysis_fixed <- lm(calcification~as.factor(pCO2),data=cal_red2)
anova(cal_analysis_fixed)

# Doesn't seem to matter, considering pCO2 as a factor and running an anova leads to a significant treatment result
# this is driven by a decrease in calcification at the high OA treatment
```

**Regression**
```{r}
## performing a regression with tank mean pCO2
cal_lm_full <- lmer(calcification~pCO2_calc + (1|PopOrigin) + (1|shelf/tank),data=cal_red2)
ranova(cal_lm_full)
#Can remove are not significant
cal_lm_Rtank <- lmer(calcification~pCO2_calc + (1|tank:shelf),data=cal_red2)
# 
summary(cal_lm_Rtank )
ranova(cal_lm_Rtank) # close but not significant effect of tank

# Final simple model (also turn calcification rate into a percentage)
cal_simple <- lm(calcification~pCO2_calc,data=cal_red2)
(cal_sum <- summary(cal_simple))
# Treatment significant
plot(cal_simple)
# Looks like some minor violations to assumptions but nothing major
```

## Calcification vs EPF Fluid pH

### Preliminary calcification plot vs. EPF carb chemistry
```{r}
cal_epf <- inner_join(cal_red2,pheno_red)

#Calcification vs EPF pH
# All points
ggplot(cal_epf,aes(x=EPF_pH,
                         y=calcification,
                         colour=as.factor(pCO2))) + 
  geom_point(size=3) +
    labs(colour="pCO2",y="Calcification(Adj)") +
  theme_bw()

# Outdated look at relationship between calcification and other carb params
# we had low sample size, which prevented statistical analysis, but there was limited correlation between calcification and other parameters

# # EPF DIC vs calcification
# ggplot(calcification_red[calcification_red$carb_chem == "Y",],aes(x=EPF_DIC,
#                          y=Pcnt_DailyChange_Exposure,
#                          colour=log(pCO2_calc))) +
#   geom_point() + geom_smooth(method="lm")
# # Trend but hard to say anything given the lack of points
# 
# # EPF Calcite Sat. vs calcification
# ggplot(calcification_red[calcification_red$carb_chem == "Y",],aes(x=EPF_Calcite_Calc,
#                          y=Pcnt_DailyChange_Exposure,
#                          colour=log(pCO2_calc))) +
#   geom_point() + geom_smooth(method="lm")

# Hard to say whats going on, but it looks like DIC is negatively correlated with calcification rate, while there seems to be a very moderate negative relationship between calcite saturation and calcification.
```

### Statistical Analysis

```{r}
#Check the impact of random factors, they didn't seem to help o they were removed
calVspH_full <- lmer(calcification ~ EPF_pH+sample_date + (1|PopOrigin) + (1|shelf/tank),data=cal_epf)
# Singular fit, checking random effects
ranova(calVspH_full)
summary(calVspH_full)
# None of the random effects were significant to I will remove first population
calVspH_randTank <- lmer(calcification ~ EPF_pH+sample_date+ (1|tank:shelf),data=cal_epf)
# Simple lm with interaction
ranova(calVspH_randTank)
summary(calVspH_randTank)

calVspH_fixed <- lm(calcification ~ EPF_pH+sample_date,data=cal_epf)
summary(calVspH_fixed)
# Model significant but no individual explanatory variables (decided to remove interaction)
# Removing interaction
calVspH_simple <- lm(calcification ~ EPF_pH,data=cal_epf)
(final.model <- summary(calVspH_simple))
plot(calVspH_simple) # No major assumption violations

# Final model has just the EPF_pH and no random effects,
# However tank effect as nearly significant so maybe want to consider putting it back, this will lead to a nonsignificant result.
```

## Final Plots 
```{r}
##### Plot 1 #####
#Mean calcification rate 
mean_cal <- aggregate(calcification*100~pCO2,cal_red2,FUN=mean)
se_cal <- aggregate(calcification*100~pCO2,cal_red2,FUN=se)
se_pco2 <- aggregate(pCO2_calc~pCO2,cal_red2,FUN=sd)
mean_cal <- data.frame(pCO2=mean_cal$pCO2,Rel_Change=mean_cal$calcification,
                       ymin=mean_cal$calcification-c(se_cal$calcification*1.96),
                       ymax=mean_cal$calcification+c(se_cal$calcification*1.96),
                         xmin=mean_cal$pCO2-c(se_pco2$pCO2_calc*1.96),
                         xmax=mean_cal$pCO2+c(se_pco2$pCO2_calc*1.96))

# ggplot version
out <- summary(lm(calcification*100~pCO2,data=cal_red2))

# p <- ggplot(cal_red2) +
#   geom_point(aes(x=pCO2_calc,y=calcification))  +
#   geom_point(data=mean_cal,aes(x=pCO2,y=Rel_Change))
# p


p <- ggplot(mean_cal,aes(x=pCO2,y=Rel_Change,shape=as.factor(pCO2),colour=as.factor(pCO2))) +
  geom_abline(slope = out$coefficients[2,1],intercept = out$coefficients[1,1]) +
  geom_hline(aes(yintercept=0),linetype="dotted") +
  geom_point(aes(size=1.5)) +
  ylim(-0.05,0.05) + 
  xlim(300,3200) +
  scale_shape_manual(values=c(16,15,17))+
  scale_colour_manual(values=c(col_perm[2],col_perm[5],col_perm[4]))+
  geom_errorbarh(aes(xmin=xmin, xmax=xmax)) + 
  geom_errorbar(aes(ymin=ymin,ymax=ymax),width=75) 

t <- p + theme_bw(base_size = 16) + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
p1<-t + ylab("Calcification (% change per day)") + xlab("pCO2 (ppm)") +
  coord_cartesian(clip = 'off') +
   geom_text(
    x = 100,
    y = 0.06,
    inherit.aes = FALSE,
    label = "A",
    check_overlap = FALSE,
    hjust = 1,
    size = 10
  ) +
  theme(legend.position="none",plot.margin = unit(c(5, 1, 1, 1), "lines")) 
p1

## Plot 2
cal_epf$pCO2_name <- "NA"
cal_epf$pCO2_name[cal_epf$pCO2 == unique(cal_epf$pCO2)[1]] <-  "Ambient"
cal_epf$pCO2_name[cal_epf$pCO2 == unique(cal_epf$pCO2)[2]] <-  "OA 900" 
cal_epf$pCO2_name[cal_epf$pCO2 == unique(cal_epf$pCO2)[3]] <-  "OA 2800"
cal_epf$pCO2_name <- factor(cal_epf$pCO2_name,levels = c("Ambient", "OA 900", "OA 2800"))

p <- ggplot(cal_epf,aes(x=EPF_pH,y=calcification*100,colour=pCO2_name)) +
  geom_point(size=4,aes(shape=pCO2_name)) + 
  ylim(-0.05,0.05) +
  scale_color_manual(values=c(col_perm[2],col_perm[5],col_perm[4])) +
  scale_shape_manual(values=c(16,15,17)) +
  geom_abline(slope = calVspH_simple$coefficients[2],
              intercept = calVspH_simple$coefficients[1]) 

t <- p + theme_bw(base_size = 16) + 
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = c(0.15,.85),
        axis.line = element_line(colour = "black")) 

p2<-t + ylab("Calcification (% change per day)") + xlab("EPF pH (NBS)") + labs(colour="pCO2 (ppm)",shape="pCO2 (ppm)") +
guides(shape = guide_legend(override.aes = list(size = 4))) +
   coord_cartesian(clip = 'off') +
   geom_text(
    x = 6.7,
    y = 0.06,
    inherit.aes = FALSE,
    label = "B",
    check_overlap = FALSE,
    hjust = 1,
    size = 10
  ) +
  theme(plot.margin = unit(c(5, 1, 1, 1), "lines"))
p2
multiplot(p1,p2,cols=2)
```

Base Plot versions (outdated)
```{r echo=FALSE}
# Base plot
# bp <- plot(mean_cal$Pcnt_DailyChange_ExposureAdj~as.numeric(as.character(mean_cal$pCO2.x)),
#            ylab=expression('Relative Calcification Rate (% day' ^{-1}*')'),
#            xlab="pCO2 (ppm)",
#            ylim = c(-0.08,0.07),xlim=c(400,3000),
#            col="black",cex=1.4,pch=16,bty="l")
# abline(h=0,lty=2)
# 
# # Fitted linear lines
# abline(lm(Pcnt_DailyChange_ExposureAdj~pCO2.x,data=calcification_red),col="black")
# 
# arrows(y0 = as.numeric(as.character(mean_cal$pCO2.x)),
#        y1 = as.numeric(as.character(mean_cal$pCO2.x)),
#        x0 = c(mean_cal$Pcnt_DailyChange_ExposureAdj - se_cal$Pcnt_DailyChange_ExposureAdj),
#        x1 = c(mean_cal$Pcnt_DailyChange_ExposureAdj + se_cal$Pcnt_DailyChange_ExposureAdj),
#        angle = 90, len = 0.05,
#        code = 3, xpd = NA, lwd = 2,
#        col="black")
# arrows(x0 = as.numeric(as.character(mean_cal$pCO2.x))- se_pco2$pCO2_calc,
#        x1 = as.numeric(as.character(mean_cal$pCO2.x))+ se_pco2$pCO2_calc,
#        y0 = c(mean_cal$Pcnt_DailyChange_ExposureAdj),
#        y1 = c(mean_cal$Pcnt_DailyChange_ExposureAdj),
#        angle = 180, len = 0.05,
#        code = 3, xpd = NA, lwd = 2,
#        col="black")

#### Plot 2 #####
# plotting time along the x-axis
# mean_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint*pCO2.x,data=calcification_red,FUN=mean)
# se_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint*pCO2.x,data=calcification_red,FUN=se)
# # Color palattee
# rbPal <- colorRampPalette(c('lightblue2','tomato'))
# calcification_red$Col2 <- rbPal(2800)[as.numeric(cut(calcification_red$pCO2_calc,breaks =2800))]
# 
# plot(calcification_red$Pcnt_DailyChange_ExposureAdj~calcification_red$EPF_pH,
#      col=calcification_red$Col2,pch=16,cex=1.4,
#      ylab=expression('Relative Calcification Rate (% day' ^{-1}*')'),
#            xlab="EPF pH (NBS)")
# abline(a=-0.50653,b=0.06607,col="Black")
# text(x=7.05,y=0.03,labels = "y = 0.067x - 0.507",pos=4)
# text(x=7.05,y=0.02,labels = expression('R'^{2}*' = 0.248'),pos=4)
# legend(y=0.055,x=6.89,title=expression('pCO'[2]*' (ppm)'),legend=seq(from=500,to=2900,length.out = 5),col=rbPal(5),pch=20,bty="n")
```

Pub plot (outdated)
```{r echo=FALSE}
# ### Final two panel plot
# par(mfrow=c(1,2))
# ##### Plot 1 #####
# #Mean calcification rate separated by time point
# cal_50 <- calcification_red[calcification_red$timepoint == 50,]
# cal_80 <- calcification_red[calcification_red$timepoint == 80,]
# mean_50_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint_fac+pCO2.x ,cal_50,FUN=mean)
# mean_80_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint_fac+pCO2.x ,cal_80,FUN=mean)
# mean_80_cal$pCO2.x <- mean_80_cal$pCO2.x + 100
# se_50_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint_fac+pCO2.x ,cal_50,FUN=se)
# se_80_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint_fac+pCO2.x ,cal_80,FUN=se)
# 
# bp <- plot(mean_50_cal$Pcnt_DailyChange_ExposureAdj~as.numeric(as.character(mean_50_cal$pCO2.x)),
#            ylab=expression('Relative Calcification Rate (% day' ^{-1}*')'),
#            xlab="pCO2 (ppm)",
#            ylim = c(-0.08,0.07),xlim=c(400,3000),
#            col="grey",cex=1.4,pch=16,bty="l")
# abline(h=0,lty=2)
# 
# points(mean_80_cal$Pcnt_DailyChange_ExposureAdj~as.numeric(as.character(mean_80_cal$pCO2.x)),col="black",
#        cex=1.4,pch=15)
# # Fitted linear lines
# abline(lm(Pcnt_DailyChange_ExposureAdj ~pCO2_calc,data=calcification_red[calcification_red$timepoint==50,]),col="grey")
# abline(lm(Pcnt_DailyChange_ExposureAdj ~pCO2_calc,data=calcification_red[calcification_red$timepoint==80,]),col="black")
# # Standard Error bars
# arrows(x0 = as.numeric(as.character(mean_50_cal$pCO2.x)),
#        x1 = as.numeric(as.character(mean_50_cal$pCO2.x)),
#        y0 = c(mean_50_cal$Pcnt_DailyChange_ExposureAdj - se_50_cal$Pcnt_DailyChange_ExposureAdj),
#        y1 = c(mean_50_cal$Pcnt_DailyChange_ExposureAdj + se_50_cal$Pcnt_DailyChange_ExposureAdj),
#        angle = 90, len = 0.05,
#        code = 3, xpd = NA, lwd = 2,
#        col="grey")
# arrows(x0 = as.numeric(as.character(mean_80_cal$pCO2.x)),
#        x1 = as.numeric(as.character(mean_80_cal$pCO2.x)),
#        y0 = c(mean_80_cal$Pcnt_DailyChange_ExposureAdj - se_80_cal$Pcnt_DailyChange_ExposureAdj),
#        y1 = c(mean_80_cal$Pcnt_DailyChange_ExposureAdj + se_80_cal$Pcnt_DailyChange_ExposureAdj),
#        angle = 90, len = 0.05,
#        code = 3, xpd = NA, lwd = 2,
#        col="black")
# # Significance bars
# segments(x0=450,y0=0.055,x1=2850,y1=0.055,col = "black")
# text(labels="**",x=c((2850-450)/2+450),y=0.06,col = "black",cex=1.8)
# segments(x0=950,y0=0.04,x1=2850,y1=0.04,col = "black")
# text(labels="**",x=c((2850-950)/2+950),y=0.045,col = "black",cex=1.8)
# # Legene
# legend(y=-0.03,x=500,title="",legend=c("Day 50","Day 80"),col=c("grey","black"),pch=c(16,15),cex=1.2,bty="n")
# 
# text("A",x=200,y=0.085,cex=2.5,xpd=TRUE)
# 
# 
# #### Plot 2 #####
# # plotting time along the x-axis
# mean_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint*pCO2.x,data=calcification_red,FUN=mean)
# se_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint*pCO2.x,data=calcification_red,FUN=se)
# # Color palattee
# rbPal <- colorRampPalette(c('lightblue2','tomato'))
# calcification_red$Col2 <- rbPal(2800)[as.numeric(cut(calcification_red$pCO2_calc,breaks =2800))]
# 
# plot(calcification_red$Pcnt_DailyChange_ExposureAdj~calcification_red$EPF_pH,
#      col=calcification_red$Col2,pch=16,cex=1.4,
#      ylab=expression('Relative Calcification Rate (% day' ^{-1}*')'),
#            xlab="EPF pH (NBS)")
# abline(a=-0.50653,b=0.06607,col="Black")
# text(x=6.91,y=0.03,labels = "y = 0.067x - 0.507",pos=4)
# text(x=6.91,y=0.025,labels = expression('R'^{2}*' = 0.248'),pos=4)
# legend(y=0.07,x=6.91,title=expression('pCO'[2]*' (ppm)'),legend=seq(from=500,to=2900,length.out = 5),col=rbPal(5),pch=20,bty="n")
# text("B",x=6.82,y=0.085,cex=2.5,xpd=TRUE)
```

## Final Thoughts 

**Final Models**

Calcification vs Env
$$Calcification_{Adj} = pCO2$$
Anova (type three)
```{r}
Anova(cal_wo81_final,type=3)
```
Regression (preferred over an ANOVA)
```{r}
sum_store
```

Calcification vs EPF pH
$$Calcification_{Adj} = EPF_{pH}$$
Regression
```{r}
summary(calVspH_final)
```

### Interpretation

**Calcification vs Environment**  
* There is a significant effect of treatment on calcification rate, which is driven by a significant decrease in calcification at the high OA treatment (2800), relative to either ambient or moderate OA. Although it would appear that there is a subtle different in the moderate OA vs. ambient conditions too, but we don't have the power to detect it.  

**Calcification vs. EPF pH**  
* Model shows there is a significant positive relationship between EPF_ph and calcification.
* Note this was not run on the other EPF carb chem parameters due to very low sample size, but from the plots it looked like DIC did not have the same relationship.

### Additional Notes  
* The choice to remove tp 81 was driven by the EPF pH. which seemed to have some measured issues (as observed in the pH under ambient being very acidic compared to all other tps). Rather than selectively keep some measures from that timepoint or using that timepoint only for calcification I decided to remove it given that that particular sampling was extra anyways and that we we used consistent individuals for all analysis.


**Second Version of the analysis redoing calcificiation estimates**
```{r eval=FALSE}
cal_v2$sample_date
cal_v2$DryWgtDate <- as.Date(paste0(substr(cal_v2$sample_date,1,4),"/",substr(cal_v2$sample_date,5,6),"/",substr(cal_v2$sample_date,7,8)))
cal_v2$date_bw2_asDate <- as.character(cal_v2$date_bw2_asDate)
cal_v2$date_bw2_asDate[cal_v2$date_bw2_asDate == "#VALUE!"] <- NA
  
cal_v2$BW2Date <- as.Date(as.character(cal_v2$date_bw2_asDate),format="%m/%d/%y")
cal_v2$DiffDryVBW2Date <- cal_v2$DryWgtDate-cal_v2$BW2Date

# subsetting data with missing values or are beyond scope of study.
cal_s <- cal_v2[!is.na(cal_v2$DiffDryVBW2Date),]
cal_s <- cal_s[cal_s$DiffDryVBW2Date < 83,]
cal_s <- cal_s[cal_s$DiffDryVBW2Date != 12,]

# Using the values from regression determine when we fit the dry weights with the bouyants 
cal_s$DryWt1Calc <- 1.87*cal_s$buoyant_start-2.74 # Dry Wgt Est. for BW1
cal_s$DryWt2Calc <- 1.87*cal_s$buoyant_2-2.74 # Dry Wgt Est. for BW2
cal_s$DryWt3Calc <- 1.87*cal_s$buoyant_3-2.74 # Dry Wgt Est. for BW3

# Baseline growth estimate from bouyant weights
cal_s$exp_calrate_basline <- (cal_s$DryWt2Calc-cal_s$DryWt1Calc)/as.numeric(cal_s$Day_BW1_BW2)
# Treatment growth rates from bouyant weights
cal_s$exp_calrate <- (cal_s$dry_end_meas-cal_s$DryWt2Calc)/as.numeric(cal_s$DiffDryVBW2Date)

## Plotting baseline growth rates
ggplot(cal_s,aes(x=as.factor(DryWgtDate),y=cal_s$exp_calrate_basline,group=as.factor(interaction(DryWgtDate,pCO2)),col=as.factor(pCO2))) + geom_boxplot()
## Plotting individuals 400 treatment ind. using final timepoint(BW3)-BW1
cal_400 <- cal_s[cal_s$pCO2 == 400,]
cal_400$exp_calrate <- (cal_400$dry_end_meas-cal_400$DryWt1Calc)/as.numeric(cal_400$Day_BW1_BW2 + cal_400$DiffDryVBW2Date)
is.na(cal_400$exp_calrate)
ggplot(cal_400,aes(x=as.factor(DryWgtDate),y=exp_calrate)) + geom_boxplot()

## Using the actual dry weight as the final weight ( Dry weight - Dry wgt. calculated from BW2)
ggplot(cal_s,aes(x=as.factor(DryWgtDate),y=cal_s$exp_calrate,group=as.factor(interaction(DryWgtDate,pCO2)),col=as.factor(pCO2))) + geom_boxplot()
# creating a model with 
lm.out <- lm(exp_calrate~DiffDryVBW2Date*as.factor(pCO2),data=cal_s)
summary(lm.out)
anova(lm.out)
# Adjusting the cal rate for oyster size (calcification rate as a % growth)
cal_s$exp_calrate_scaled <- cal_s$exp_calrate/cal_s$dry_end_meas
ggplot(cal_s,aes(x=as.factor(DryWgtDate),y=cal_s$exp_calrate_scaled,group=as.factor(interaction(DryWgtDate,pCO2)),col=as.factor(pCO2))) + geom_boxplot()


## Using the dry weights converted from bouyant wgt
cal_s$exp_calrate_fromBouyant <- (cal_s$DryWt3Calc-cal_s$DryWt2Calc)/as.numeric(cal_s$Day_BW_BW3)
ggplot(cal_s,aes(x=as.factor(pCO2),y=cal_s$exp_calrate_fromBouyant,group=as.factor(interaction(pCO2)),col=as.factor(pCO2))) + geom_boxplot()


```
