---
title: "AE17 Calcification Full Workup"
author: "adowneywall"
date: "8/21/2019"
output: github_document
editor_options:     
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(lmerTest)
library(car)
source("/home/downeyam/R/basicR_functions.R")
library(RColorBrewer)
pal <- brewer.pal(n = 12, name = 'Paired')
col_perm <- c(pal[1:2],pal[5:6],pal[12])
```

### Data
```{r}
# Dataset with all timepoints
calcification <- readRDS("/home/downeyam/Github/2017OAExp_Oysters/input_files/Phenotype/SummaryCalcificationData.RData")
# Dataset without tp 81
calcification_red <- readRDS("/home/downeyam/Github/2017OAExp_Oysters/input_files/Phenotype/SummaryCalcificationData_without81.RData")
cal_v2<-read.csv("/home/downeyam/Github/2017OAExp_Oysters/input_files/Phenotype/EPF/AE17_Exp_2017_CalcificationInfo_20190903_Final.csv")
```

### Primary Variables 

**Fixed Effects**

* Treatment 
  * pCO2_fac = Treating pCO2 like a factor
  * pCO2_calc = Treating pCO2 as continuous variable using tank mean values

* Time
  * Timepoint_fac =  Two (or three) timepoints as factors
  * timepoint = Two timepoint as continous variable

**Random Effects**

* Pop : 3 sites -  roughly 1/3 oysters per site
* Shelf : 6 shelfs (nested in treatment) - 12-18 oysters per shelf
* Tank : 3 tanks per shelf (nested) - 2 -3 oysters per tank


### Tables

**All timepoints**
```{r}
# Table
table(calcification$timepoint,calcification$pCO2_fac)
# Total number of samples
nrow(calcification)
```

**With final timepoint removed**  
Rationale explained in notes at end
```{r}
# Table
table(calcification_red$timepoint,calcification_red$pCO2_fac)
# Total number of samples
nrow(calcification_red)
```

## Preliminary Plots with Calcification Data
```{r}
### Plotting Calcification Data
## Calcification Adj -- with Time -- Full Data
gg <- ggplot(calcification,aes(x=pCO2.x,y=Pcnt_DailyChange_ExposureAdj,
                         group=interaction(pCO2.x,timepoint),
                         colour=interaction(pCO2.x,timepoint))) 
gg + geom_boxplot() +
  labs(colour="pCO2.Timepoint",x="pCO2",title="All timepoints")
## Calcification Adj -- with Time --sans day 81 Data
ggplot(calcification_red,aes(x=pCO2.x,y=Pcnt_DailyChange_ExposureAdj,
                         group=interaction(pCO2.x,timepoint),
                         colour=interaction(pCO2.x,timepoint))) + 
  geom_boxplot() + labs(colour="pCO2.Timepoint",title="Without tp 81")
## Calcification Adj -- without Time -- Full Data
ggplot(calcification,aes(x=pCO2.x,y=Pcnt_DailyChange_ExposureAdj,
                         group=(pCO2.x))) + geom_boxplot() +
  labs(x="pCO2",title="All timepoints")
## Calcification Adj -- without Time -- sans 81 Data
ggplot(calcification_red,aes(x=pCO2.x,y=Pcnt_DailyChange_ExposureAdj,
                         group=(pCO2.x))) + geom_boxplot() +
  labs(x="pCO2",title="Without tp 81")
## Calcification -- without Time -- Full Data
ggplot(calcification,aes(x=pCO2.x,y=Pcnt_DailyChange_Exposure,
                         group=(pCO2.x))) + geom_boxplot()
## Calcification -- without Time -- sans 81 Data
ggplot(calcification_red,aes(x=pCO2.x,y=Pcnt_DailyChange_Exposure,
                         group=(pCO2.x))) + geom_boxplot()
ggplot(calcification_red,aes(x=pCO2_calc,y=Pcnt_DailyChange_Exposure,
                         group=(pCO2.x))) + geom_point()
# Briefly check out potential site specific effects
boxplot(calcification_red$Pcnt_DailyChange_Exposure~calcification_red$PopOrigin*calcification_red$pCO2_fac)
# No obvious site specific effevt
```

## Calcification vs Environment and Time - Statistical Analysis

**ANOVA**
```{r}
# Full model- All points
cal_full <- lmer(Pcnt_DailyChange_ExposureAdj~timepoint_fac*pCO2_fac + (1|PopOrigin) + (1|shelf/tank),data=calcification)
# Singular fit so we need to simplfy
anova(cal_full) # No significance
ranova(cal_full)
#None are significant / Remove pop and shelf and rerun

# Reducing model - keeping fixed factors and one random effect
cal_full_2 <- lmer(Pcnt_DailyChange_ExposureAdj~timepoint_fac+pCO2_fac + (1|shelf),data=calcification)
ranova(cal_full_2)
# Can't include due to singular fit, and its not significant. 

# Fixed effects only
cal_analysis_fixed <- lm(Pcnt_DailyChange_ExposureAdj~pCO2_calc*timepoint,data=calcification)
step(cal_analysis_fixed)
# Step function indicates a mode that includes pco2 and time but not the interaction is the best approach
cal_analysis_final <- lm(Pcnt_DailyChange_ExposureAdj~pCO2_calc+timepoint,data=calcification)
anova(cal_analysis_final)

### Full Model - With 81 removed
cal_wo81_full <- lmer(Pcnt_DailyChange_ExposureAdj~timepoint_fac*pCO2_calc + (1|PopOrigin) + (1|shelf/tank),data=calcification_red)
anova(cal_wo81_full)
ranova(cal_wo81_full)
#Similar results to full data so moving directly to fixed model
cal_wo81_fixed <- lm(Pcnt_DailyChange_ExposureAdj~pCO2_calc*timepoint,data=calcification_red)
step(cal_wo81_fixed)
# Final model is the two fixed factors but not the interaction
cal_wo81_final <- lm(Pcnt_DailyChange_ExposureAdj~pCO2_calc+timepoint,data=calcification_red)
# Type 3
Anova(cal_wo81_final,type=2)
# Same general interpretation as all timepoints. Treatment is significant and time doesn't matter.

# Plot post hoc testing
plot(TukeyHSD(aov(Pcnt_DailyChange_ExposureAdj~pCO2_fac+timepoint_fac, data=calcification_red)))
#Post hoc reveals that its the 2800 being significantly differen't thatn either 900 or 400 that is driving the pattern
```

**Regression**
```{r}
## performing a regression with tank mean pCO2
cal_analysis2 <- lmer(Pcnt_DailyChange_ExposureAdj~pCO2_calc*timepoint_fac+(1|PopOrigin)+(1|shelf),data=calcification_red)
ranova(cal_analysis2)
#Can remove are not significant
cal_interaction <- lm(Pcnt_DailyChange_ExposureAdj~pCO2_calc*timepoint_fac,data=calcification_red)
# Interaction not significant, so it will be dropped
summary(cal_interaction)
cal_simple <- lm(Pcnt_DailyChange_ExposureAdj~pCO2_calc+timepoint_fac,data=calcification_red)
(cal_sum <- summary(cal_simple))
# Treatment but not time significant. I guess I will only plot a fitted line based on the simplest model with only treatment.
cal_pCO2 <- lm(Pcnt_DailyChange_ExposureAdj~pCO2_calc,data=calcification_red)
(sum_store <- summary(cal_pCO2))
# A few slightly wonky points, but no major violations
```

## Calcification vs EPF Fluid pH

### Preliminary calcification plot vs. EPF carb chemistry
```{r}
#Calcification vs pH
# All points
ggplot(calcification,aes(x=EPF_pH,
                         y=Pcnt_DailyChange_ExposureAdj,
                         colour=pCO2_calc)) +
  geom_point() + geom_smooth(method="lm")+labs(colour="pCO2",y="Calcification(Adj)")
# Removing tp 81 because points were wonky
ggplot(calcification_red,aes(x=EPF_pH,
                         y=Pcnt_DailyChange_ExposureAdj,
                         colour=pCO2_calc)) +
  geom_point() + geom_smooth(method="lm",colour="black") +
  labs(colour="pCO2",y="Calcification(Adj)") +
  theme_bw()

# EPF DIC vs calcification
ggplot(calcification_red[calcification_red$carb_chem == "Y",],aes(x=EPF_DIC,
                         y=Pcnt_DailyChange_Exposure,
                         colour=log(pCO2_calc))) +
  geom_point() + geom_smooth(method="lm")
# Trend but hard to say anything given the lack of points

# EPF Calcite Sat. vs calcification
ggplot(calcification_red[calcification_red$carb_chem == "Y",],aes(x=EPF_Calcite_Calc,
                         y=Pcnt_DailyChange_Exposure,
                         colour=log(pCO2_calc))) +
  geom_point() + geom_smooth(method="lm")

# Hard to say whats going on, but it looks like DIC is negatively correlated with calcification rate, while there seems to be a very moderate negative relationship between calcite saturation and calcification.
```

### Statistical Analysis

```{r}
#Check the impact of random factors, they didn't seem to help o they were removed
calVspH_full <- lmer(Pcnt_DailyChange_ExposureAdj ~ EPF_pH * timepoint + (1|PopOrigin) + (1|shelf/tank),data=calcification_red)
# singular fit
ranova(calVspH_full)
summary(calVspH_full)
# None of the random effects were significant to I will remove
# Simple lm with interaction
calVspH_fixed_inter <- lm(Pcnt_DailyChange_ExposureAdj ~ EPF_pH * timepoint ,data=calcification_red)
summary(calVspH_fixed_inter)
# Model significant but no individual explanatory variables (decided to remove interaction)
# Removing interaction
calVspH_fixed <- lmer(Pcnt_DailyChange_ExposureAdj ~ EPF_pH + (1|tank:shelf) ,data=calcification_red)
(calVpH_fized_out <- summary(calVspH_fixed))
ranova(calVspH_fixed)
# Treatment significant but not time
# Little evidence of a tank effect so it's removed
calVspH_final <- lm(Pcnt_DailyChange_ExposureAdj ~ EPF_pH ,data=calcification_red)
(calVpH_fized_out <- summary(calVspH_final))
# With out without random effect of tank the relationship between EPF pH and calcification is significant

```

## Final Plots 
```{r}
##### Plot 1 #####
#Mean calcification rate 
mean_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~pCO2.x ,calcification_red,FUN=mean)
se_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~pCO2.x ,calcification_red,FUN=se)
se_pco2 <- aggregate(pCO2_calc~pCO2.x ,calcification_red,FUN=sd)
mean_cal <- data.frame(pCO2=mean_cal$pCO2.x,Rel_Change=mean_cal$Pcnt_DailyChange_ExposureAdj,
                         ymin=mean_cal$Pcnt_DailyChange_ExposureAdj-c(se_cal$Pcnt_DailyChange_ExposureAdj*1.96),
                         ymax=mean_cal$Pcnt_DailyChange_ExposureAdj+c(se_cal$Pcnt_DailyChange_ExposureAdj*1.96),
                         xmin=mean_cal$pCO2.x-c(se_pco2$pCO2_calc*1.96),
                         xmax=mean_cal$pCO2.x+c(se_pco2$pCO2_calc*1.96))

# ggplot version
out <- summary(lm(Pcnt_DailyChange_ExposureAdj~pCO2.x,data=calcification_red))

p <- ggplot(mean_cal,aes(x=pCO2,y=Rel_Change)) + geom_hline(aes(yintercept=0),linetype="dotted") +
  geom_point(aes(size=6,shape=as.factor(pCO2))) + ylim(-0.085,0.08) + xlim(300,3200) + 
  scale_shape_manual(values=c(16,15,17))+
  geom_errorbarh(aes(xmin=xmin, xmax=xmax)) + 
  geom_errorbar(aes(ymin=ymin,ymax=ymax),width=75) +
  geom_abline(slope = out$coefficients[2,1],intercept = out$coefficients[1,1]) 
t <- p + theme_bw(base_size = 16) + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
p1<-t + ylab("Calcification (% change per day)") + xlab("pCO2 (ppm)") +
  coord_cartesian(clip = 'off') +
   geom_text(
    x = 100,
    y = 0.10,
    inherit.aes = FALSE,
    label = "A",
    check_overlap = FALSE,
    hjust = 1,
    size = 10
  ) +
  theme(legend.position="none",plot.margin = unit(c(5, 1, 1, 1), "lines")) 
p1

## Plot 2
calcification_red$pCO2_name <- "NA"
calcification_red$pCO2_name[calcification_red$pCO2_fac == unique(calcification_red$pCO2_fac)[1]] <-  "Ambient"
calcification_red$pCO2_name[calcification_red$pCO2_fac == unique(calcification_red$pCO2_fac)[2]] <-  "OA 900" 
calcification_red$pCO2_name[calcification_red$pCO2_fac == unique(calcification_red$pCO2_fac)[3]] <-  "OA 2800"
calcification_red$pCO2_name <- as.factor(calcification_red$pCO2_name)
levels(calcification_red$pCO2_name) <- levels(calcification_red$pCO2_name)[c(1,3,2)]

p <- ggplot(calcification_red,aes(x=EPF_pH,y=Pcnt_DailyChange_ExposureAdj,colour=pCO2_name)) +
  geom_point(size=4,aes(shape=pCO2_name)) + ylim(-0.085,0.08) +
  scale_color_manual(values=c(col_perm[2],col_perm[5],col_perm[4])) +
  scale_shape_manual(values=c(16,15,17))+
  geom_abline(slope = calVpH_fized_out$coefficients[2,1],intercept = calVpH_fized_out$coefficients[1,1]) 

t <- p + theme_bw(base_size = 16) + 
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = c(0.15,.85),
        axis.line = element_line(colour = "black")) 

p2<-t + ylab("Calcification (% change per day)") + xlab("EPF pH (NBS)") + labs(colour="pCO2 (ppm)",shape="pCO2 (ppm)") +
guides(shape = guide_legend(override.aes = list(size = 4))) +
   coord_cartesian(clip = 'off') +
   geom_text(
    x = 6.85,
    y = 0.10,
    inherit.aes = FALSE,
    label = "B",
    check_overlap = FALSE,
    hjust = 1,
    size = 10
  ) +
  theme(plot.margin = unit(c(5, 1, 1, 1), "lines"))
p2
multiplot(p1, p2,cols=2)
```

Base Plot versions (outdated)
```{r echo=FALSE}
# Base plot
# bp <- plot(mean_cal$Pcnt_DailyChange_ExposureAdj~as.numeric(as.character(mean_cal$pCO2.x)),
#            ylab=expression('Relative Calcification Rate (% day' ^{-1}*')'),
#            xlab="pCO2 (ppm)",
#            ylim = c(-0.08,0.07),xlim=c(400,3000),
#            col="black",cex=1.4,pch=16,bty="l")
# abline(h=0,lty=2)
# 
# # Fitted linear lines
# abline(lm(Pcnt_DailyChange_ExposureAdj~pCO2.x,data=calcification_red),col="black")
# 
# arrows(y0 = as.numeric(as.character(mean_cal$pCO2.x)),
#        y1 = as.numeric(as.character(mean_cal$pCO2.x)),
#        x0 = c(mean_cal$Pcnt_DailyChange_ExposureAdj - se_cal$Pcnt_DailyChange_ExposureAdj),
#        x1 = c(mean_cal$Pcnt_DailyChange_ExposureAdj + se_cal$Pcnt_DailyChange_ExposureAdj),
#        angle = 90, len = 0.05,
#        code = 3, xpd = NA, lwd = 2,
#        col="black")
# arrows(x0 = as.numeric(as.character(mean_cal$pCO2.x))- se_pco2$pCO2_calc,
#        x1 = as.numeric(as.character(mean_cal$pCO2.x))+ se_pco2$pCO2_calc,
#        y0 = c(mean_cal$Pcnt_DailyChange_ExposureAdj),
#        y1 = c(mean_cal$Pcnt_DailyChange_ExposureAdj),
#        angle = 180, len = 0.05,
#        code = 3, xpd = NA, lwd = 2,
#        col="black")

#### Plot 2 #####
# plotting time along the x-axis
# mean_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint*pCO2.x,data=calcification_red,FUN=mean)
# se_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint*pCO2.x,data=calcification_red,FUN=se)
# # Color palattee
# rbPal <- colorRampPalette(c('lightblue2','tomato'))
# calcification_red$Col2 <- rbPal(2800)[as.numeric(cut(calcification_red$pCO2_calc,breaks =2800))]
# 
# plot(calcification_red$Pcnt_DailyChange_ExposureAdj~calcification_red$EPF_pH,
#      col=calcification_red$Col2,pch=16,cex=1.4,
#      ylab=expression('Relative Calcification Rate (% day' ^{-1}*')'),
#            xlab="EPF pH (NBS)")
# abline(a=-0.50653,b=0.06607,col="Black")
# text(x=7.05,y=0.03,labels = "y = 0.067x - 0.507",pos=4)
# text(x=7.05,y=0.02,labels = expression('R'^{2}*' = 0.248'),pos=4)
# legend(y=0.055,x=6.89,title=expression('pCO'[2]*' (ppm)'),legend=seq(from=500,to=2900,length.out = 5),col=rbPal(5),pch=20,bty="n")
```

Pub plot (outdated)
```{r echo=FALSE}
# ### Final two panel plot
# par(mfrow=c(1,2))
# ##### Plot 1 #####
# #Mean calcification rate separated by time point
# cal_50 <- calcification_red[calcification_red$timepoint == 50,]
# cal_80 <- calcification_red[calcification_red$timepoint == 80,]
# mean_50_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint_fac+pCO2.x ,cal_50,FUN=mean)
# mean_80_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint_fac+pCO2.x ,cal_80,FUN=mean)
# mean_80_cal$pCO2.x <- mean_80_cal$pCO2.x + 100
# se_50_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint_fac+pCO2.x ,cal_50,FUN=se)
# se_80_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint_fac+pCO2.x ,cal_80,FUN=se)
# 
# bp <- plot(mean_50_cal$Pcnt_DailyChange_ExposureAdj~as.numeric(as.character(mean_50_cal$pCO2.x)),
#            ylab=expression('Relative Calcification Rate (% day' ^{-1}*')'),
#            xlab="pCO2 (ppm)",
#            ylim = c(-0.08,0.07),xlim=c(400,3000),
#            col="grey",cex=1.4,pch=16,bty="l")
# abline(h=0,lty=2)
# 
# points(mean_80_cal$Pcnt_DailyChange_ExposureAdj~as.numeric(as.character(mean_80_cal$pCO2.x)),col="black",
#        cex=1.4,pch=15)
# # Fitted linear lines
# abline(lm(Pcnt_DailyChange_ExposureAdj ~pCO2_calc,data=calcification_red[calcification_red$timepoint==50,]),col="grey")
# abline(lm(Pcnt_DailyChange_ExposureAdj ~pCO2_calc,data=calcification_red[calcification_red$timepoint==80,]),col="black")
# # Standard Error bars
# arrows(x0 = as.numeric(as.character(mean_50_cal$pCO2.x)),
#        x1 = as.numeric(as.character(mean_50_cal$pCO2.x)),
#        y0 = c(mean_50_cal$Pcnt_DailyChange_ExposureAdj - se_50_cal$Pcnt_DailyChange_ExposureAdj),
#        y1 = c(mean_50_cal$Pcnt_DailyChange_ExposureAdj + se_50_cal$Pcnt_DailyChange_ExposureAdj),
#        angle = 90, len = 0.05,
#        code = 3, xpd = NA, lwd = 2,
#        col="grey")
# arrows(x0 = as.numeric(as.character(mean_80_cal$pCO2.x)),
#        x1 = as.numeric(as.character(mean_80_cal$pCO2.x)),
#        y0 = c(mean_80_cal$Pcnt_DailyChange_ExposureAdj - se_80_cal$Pcnt_DailyChange_ExposureAdj),
#        y1 = c(mean_80_cal$Pcnt_DailyChange_ExposureAdj + se_80_cal$Pcnt_DailyChange_ExposureAdj),
#        angle = 90, len = 0.05,
#        code = 3, xpd = NA, lwd = 2,
#        col="black")
# # Significance bars
# segments(x0=450,y0=0.055,x1=2850,y1=0.055,col = "black")
# text(labels="**",x=c((2850-450)/2+450),y=0.06,col = "black",cex=1.8)
# segments(x0=950,y0=0.04,x1=2850,y1=0.04,col = "black")
# text(labels="**",x=c((2850-950)/2+950),y=0.045,col = "black",cex=1.8)
# # Legene
# legend(y=-0.03,x=500,title="",legend=c("Day 50","Day 80"),col=c("grey","black"),pch=c(16,15),cex=1.2,bty="n")
# 
# text("A",x=200,y=0.085,cex=2.5,xpd=TRUE)
# 
# 
# #### Plot 2 #####
# # plotting time along the x-axis
# mean_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint*pCO2.x,data=calcification_red,FUN=mean)
# se_cal <- aggregate(Pcnt_DailyChange_ExposureAdj~timepoint*pCO2.x,data=calcification_red,FUN=se)
# # Color palattee
# rbPal <- colorRampPalette(c('lightblue2','tomato'))
# calcification_red$Col2 <- rbPal(2800)[as.numeric(cut(calcification_red$pCO2_calc,breaks =2800))]
# 
# plot(calcification_red$Pcnt_DailyChange_ExposureAdj~calcification_red$EPF_pH,
#      col=calcification_red$Col2,pch=16,cex=1.4,
#      ylab=expression('Relative Calcification Rate (% day' ^{-1}*')'),
#            xlab="EPF pH (NBS)")
# abline(a=-0.50653,b=0.06607,col="Black")
# text(x=6.91,y=0.03,labels = "y = 0.067x - 0.507",pos=4)
# text(x=6.91,y=0.025,labels = expression('R'^{2}*' = 0.248'),pos=4)
# legend(y=0.07,x=6.91,title=expression('pCO'[2]*' (ppm)'),legend=seq(from=500,to=2900,length.out = 5),col=rbPal(5),pch=20,bty="n")
# text("B",x=6.82,y=0.085,cex=2.5,xpd=TRUE)
```

## Final Thoughts 

**Final Models**

Calcification vs Env
$$Calcification_{Adj} = pCO2$$
Anova (type three)
```{r}
Anova(cal_wo81_final,type=3)
```
Regression (preferred over an ANOVA)
```{r}
sum_store
```

Calcification vs EPF pH
$$Calcification_{Adj} = EPF_{pH}$$
Regression
```{r}
summary(calVspH_final)
```

### Interpretation

**Calcification vs Environment**  
* There is a significant effect of treatment on calcification rate, which is driven by a significant decrease in calcification at the high OA treatment (2800), relative to either ambient or moderate OA. Although it would appear that there is a subtle different in the moderate OA vs. ambient conditions too, but we don't have the power to detect it.  

**Calcification vs. EPF pH**  
* Model shows there is a significant positive relationship between EPF_ph and calcification.
* Note this was not run on the other EPF carb chem parameters due to very low sample size, but from the plots it looked like DIC did not have the same relationship.

### Additional Notes  
* The choice to remove tp 81 was driven by the EPF pH. which seemed to have some measured issues (as observed in the pH under ambient being very acidic compared to all other tps). Rather than selectively keep some measures from that timepoint or using that timepoint only for calcification I decided to remove it given that that particular sampling was extra anyways and that we we used consistent individuals for all analysis.


**Second Version of the analysis redoing calcificiation estimates**
```{r eval=FALSE}
cal_v2$sample_date
cal_v2$DryWgtDate <- as.Date(paste0(substr(cal_v2$sample_date,1,4),"/",substr(cal_v2$sample_date,5,6),"/",substr(cal_v2$sample_date,7,8)))
cal_v2$date_bw2_asDate <- as.character(cal_v2$date_bw2_asDate)
cal_v2$date_bw2_asDate[cal_v2$date_bw2_asDate == "#VALUE!"] <- NA
  
cal_v2$BW2Date <- as.Date(as.character(cal_v2$date_bw2_asDate),format="%m/%d/%y")
cal_v2$DiffDryVBW2Date <- cal_v2$DryWgtDate-cal_v2$BW2Date

# subsetting data with missing values or are beyond scope of study.
cal_s <- cal_v2[!is.na(cal_v2$DiffDryVBW2Date),]
cal_s <- cal_s[cal_s$DiffDryVBW2Date < 83,]
cal_s <- cal_s[cal_s$DiffDryVBW2Date != 12,]

# Using the values from regression determine when we fit the dry weights with the bouyants 
cal_s$DryWt1Calc <- 1.87*cal_s$buoyant_start-2.74 # Dry Wgt Est. for BW1
cal_s$DryWt2Calc <- 1.87*cal_s$buoyant_2-2.74 # Dry Wgt Est. for BW2
cal_s$DryWt3Calc <- 1.87*cal_s$buoyant_3-2.74 # Dry Wgt Est. for BW3

# Baseline growth estimate from bouyant weights
cal_s$exp_calrate_basline <- (cal_s$DryWt2Calc-cal_s$DryWt1Calc)/as.numeric(cal_s$Day_BW1_BW2)
# Treatment growth rates from bouyant weights
cal_s$exp_calrate <- (cal_s$dry_end_meas-cal_s$DryWt2Calc)/as.numeric(cal_s$DiffDryVBW2Date)

## Plotting baseline growth rates
ggplot(cal_s,aes(x=as.factor(DryWgtDate),y=cal_s$exp_calrate_basline,group=as.factor(interaction(DryWgtDate,pCO2)),col=as.factor(pCO2))) + geom_boxplot()
## Plotting individuals 400 treatment ind. using final timepoint(BW3)-BW1
cal_400 <- cal_s[cal_s$pCO2 == 400,]
cal_400$exp_calrate <- (cal_400$dry_end_meas-cal_400$DryWt1Calc)/as.numeric(cal_400$Day_BW1_BW2 + cal_400$DiffDryVBW2Date)
is.na(cal_400$exp_calrate)
ggplot(cal_400,aes(x=as.factor(DryWgtDate),y=exp_calrate)) + geom_boxplot()

## Using the actual dry weight as the final weight ( Dry weight - Dry wgt. calculated from BW2)
ggplot(cal_s,aes(x=as.factor(DryWgtDate),y=cal_s$exp_calrate,group=as.factor(interaction(DryWgtDate,pCO2)),col=as.factor(pCO2))) + geom_boxplot()
# creating a model with 
lm.out <- lm(exp_calrate~DiffDryVBW2Date*as.factor(pCO2),data=cal_s)
summary(lm.out)
anova(lm.out)
# Adjusting the cal rate for oyster size (calcification rate as a % growth)
cal_s$exp_calrate_scaled <- cal_s$exp_calrate/cal_s$dry_end_meas
ggplot(cal_s,aes(x=as.factor(DryWgtDate),y=cal_s$exp_calrate_scaled,group=as.factor(interaction(DryWgtDate,pCO2)),col=as.factor(pCO2))) + geom_boxplot()


## Using the dry weights converted from bouyant wgt
cal_s$exp_calrate_fromBouyant <- (cal_s$DryWt3Calc-cal_s$DryWt2Calc)/as.numeric(cal_s$Day_BW_BW3)
ggplot(cal_s,aes(x=as.factor(pCO2),y=cal_s$exp_calrate_fromBouyant,group=as.factor(interaction(pCO2)),col=as.factor(pCO2))) + geom_boxplot()


```
