---
title: "CV17_RNA_WGCNA"
author: "adowneywall"
date: "December 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(WGCNA)
knitr::opts_knit$set(root.dir = "/home/downeyam/Github/2017OAExp_Oysters/")
# The following setting is important, do not omit.
options(stringsAsFactors = FALSE)
enableWGCNAThreads()
```

```{r data_read, include=TRUE}
matData <- read.csv("results/finalNormalizedExpMatrix.csv")
matRowNames <- matData$X
matData <- matData[,c(2:25)]
row.names(matData) <- matRowNames
matData_T <- t(matData)
```

Check normalized and scaled gene data to confirm no major missing values
```{r gene_test}
gsg <-  goodSamplesGenes(matData_T, verbose = 3);
gsg$allOK
```

Next we cluster the samples (in contrast to clustering genes that will come later) to see if there are any obvious outliers.
```{r}
sampleTree <-  hclust(dist(matData_T), method = "average");
# Plot the sample tree: Open a graphic output window of size 12 by 9 inches
# The user should change the dimensions if the window is too large or too small.
sizeGrWindow(12,9)
#pdf(file = "Plots/sampleClustering.pdf", width = 12, height = 9);
par(cex = 0.6);
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5, 
     cex.axis = 1.5, cex.main = 2)

```

No evidence of outliers.

## Trait Data

We now read in the trait data and match the samples for which they were measured to the expression samples.
```{r}
traitData <-   read.table("input_files/metadata_cvirginica_rna_meta.txt",header = T)
dim(traitData)
names(traitData)

# remove columns that hold information we do not need.
allTraits <- subset(traitData,select=c("sample_name","sample_index","treatment","timepoint","shelf","tank","population","lane","dry_wgtcorr","epf_pH"))

Sample_ID <-  rownames(matData_T)
traitRows <-  match(Sample_ID, allTraits$sample_name)
datTraits <-  allTraits[traitRows, -1]
rownames(datTraits) <-  allTraits[traitRows, 1]

collectGarbage()
```


```{r}
# Re-cluster samples
sampleTree2 <-  hclust(dist(matData_T), method = "average")
# Convert traits to a color representation: white means low, red means high, grey means missing entry
traitColors <-  numbers2colors(datTraits, signed = FALSE);
# Plot the sample dendrogram and the colors underneath.
plotDendroAndColors(sampleTree2, traitColors,
                    groupLabels = names(datTraits), 
                    main = "Sample dendrogram and trait heatmap")
```

**STEPS LEFT OVER FROM TUTORIAL EDIT** 
```{r}
save(matData_T, datTraits, file = "/home/downeyam/Github/2017OAExp_Oysters/results/WGCNAdataInput.RData")
```

```{r}
# Load the data saved in the first part
lnames = load(file = "/home/downeyam/Github/2017OAExp_Oysters/results/WGCNAdataInput.RData");
#The variable lnames contains the names of loaded variables.
lnames
```

### Step-by-step network construction and module detection  

#### Choosing the soft-thresholding power: analysis of network topology  
Constructing a weighted gene network entails the choice of the soft thresholding power β to which co-expression similarity is raised to calculate adjacency. The authors of have proposed to choose the soft thresholding power based on the criterion of approximate scale-free topology. We refer the reader to that work for more details; here we illustrate the use of the function pickSoftThreshold that performs the analysis of network topology and aids the user in choosing a proper soft-thresholding power. The user chooses a set of candidate powers (the function provides suitable default values), and the function returns a set of network indices that should be inspected, for example as follows:  

```{r}
# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to=20, by=2))
# Call the network topology analysis function
sft = pickSoftThreshold(matData_T, powerVector = powers, verbose = 5)
# Plot the results:
sizeGrWindow(9, 5)
par(mfrow = c(1,2));
cex1 = 0.9;
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red");
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```





